---
title: "Figure 2 plots"
output: github_document
date: "2026-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(tidyr)
library(dplyr)
library(Matrix)
library(stringr)
library(ggrepel)
library(readr)
library(cowplot)
library(EnhancedVolcano)
library(MAGeCKFlute)

options(future.globals.maxSize = 10 * 1024^3)
```



Figures 2A,B,D  illustrator files

The gene summary can be found in Table S2 in the supplemental tables. It was generated by running Mageck on the counts table as described in the methods with the code reproduced here (command-line code)

1.Trim all of the Reads so that just the 20bp of the gRNA is still there: cutadapt -g GGACGAAACACCG -o combined_trim5.fastq combined.fastq
cutadapt -a GTTTCAGAGCT -o combined.fastq_trim.fastq combined_trim5.fastq

2.Run Mageck count on all the reads: 
mageck count -l library.csv -n test --sample-label Top1,Top2,Bottom1,Bottom2  --fastq Top1.fastq Top2.fastq bottom1.fastq bottom2.fastq

3.Run Mageck test on the counts: 
mageck test -k test.count.txt -t Top1,Top2 -c Bottom1,Bottom2 -n TvB



```{r Fig2C, dpi=300, fig.width=10, fig.height=10, dev='png'}
TvB.gene_summary <- read.delim("gene_summary.txt")
TvB.gene_summary$smaller_value <- pmin(TvB.gene_summary$neg.p.value, TvB.gene_summary$pos.p.value)
TvB.gene_summary$smaller_rank <- pmin(TvB.gene_summary$neg.rank, TvB.gene_summary$pos.rank)
TvB.gene_summary$smaller_rank <- (TvB.gene_summary$smaller_rank)/18939
genes_lab <- c(
  TvB.gene_summary$id[order(TvB.gene_summary$neg.rank)][1:25],
  TvB.gene_summary$id[order(TvB.gene_summary$pos.rank)][1:25]
)
p1 <- EnhancedVolcano(TvB.gene_summary,TvB.gene_summary$id,"neg.lfc","smaller_value",drawConnectors = T,arrowheads = F,selectLab = genes_lab,xlim = c(-6.5,6.5),ylim = c(0,7),FCcutoff = 1,pCutoff = 0.05,title = NULL,subtitle = NULL)

plot(p1)
```


The sgRNA summary is generated alongside the gene summary using the MaGeck code above
```{r Fig2E, dpi=300, fig.width=10, fig.height=10, dev='png'}
TvB.sgrna_summary <- read.delim("sgrna_summary.txt")

p3 = sgRankView(TvB.sgrna_summary,gene =c("TSC1","TSC2","ITFG2","MAP4K5","C12orf66","TTI2","SUGT1","SEM1","RPTOR","GFI1B"),top=0,bottom = 0)

plot(p3)
```

The same plots for the supplement can be generated using the Mageck code, but only running one replicate at a time
```{r FigS2B, dpi=300, fig.width=10, fig.height=10, dev='png'}
TvB_1 <- read.delim("sgrna_summary_rep1.txt")
TvB_2 <- read.delim("sgrna_summary_rep1.txt")

p1 = sgRankView(TvB_1,gene =c("GFI1B","RPTOR","SOS1","SUGT1","TTI1","C12orf66","MAP4K5","ITFG2","ERCC5","TSC1"),top=0,bottom = 0)
p2 = sgRankView(TvB_2,gene =c("GFI1B","RPTOR","SOS1","SUGT1","TTI1","C12orf66","MAP4K5","ITFG2","ERCC5","TSC1"),top=0,bottom = 0)


plot_grid(p1,p2)
```