---
title: "Fig4"
output: html_document
date: "2026-06-11"
---
---
title: "Figure 4 plots"
output: github_document
date: "2026-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(tidyr)
library(dplyr)
library(Matrix)
library(stringr)
library(ggrepel)
library(readr)
library(cowplot)
library(ggplot2)
library(tibble)


options(future.globals.maxSize = 10 * 1024^3)
```



The first steps for this part are to model. In the section below is the code used to produce the OLS, the Lasso and the Ridge model used and tested in the paper. Note that producing the correlations for pRPS6 and each RNA takes some time to run. 


```{r Modeling}

#### OLS model using Glossary Data and Applying to "Test" dataset
### First load in "Glossary", "Test" and "GWPS_mTOR" datasets
rna.mtx <- GetAssayData(Glossary,assay = "RNA",layer = "data")
adt.mtx <- GetAssayData(Glossary,assay = "ADT",layer = "data")

rna.correlations <- list()

adt_list <- "pRPS6"
for (adt in adt_list) {
  c <- list()
  print(paste("On ADT:", adt))
  for (gene in rownames(rna.mtx)) {
    r <- cor(x = adt.mtx[adt,], y = rna.mtx[gene,], method = "pearson")
    c[[gene]] <- r
  }
  rna.correlations[[adt]] <- c
}

# --- 2. Rank genes by pRPS6 correlation, select top/bottom ---
pRPS6_cors <- sort(
  unlist(rna.correlations$pRPS6),
  decreasing = TRUE,
  na.last = NA       # drops NAs automatically
)

n_genes    <- length(pRPS6_cors)
pRPS6_top  <- names(pRPS6_cors)[1:1000]
pRPS6_bot  <- names(pRPS6_cors)[(n_genes - 999):n_genes]  # bottom 1000
pRPS6_pred <- c(pRPS6_top, pRPS6_bot)

# --- 3. Filter to genes present in GWPS ---
GWPS_genes  <- rownames(GetAssayData(GWPS_mTOR@assays$RNA))
overlap     <- intersect(pRPS6_pred, GWPS_genes)

# --- 4. Build training data (Glossary) and fit OLS model ---
rna_glossary   <- as.data.frame(t(GetAssayData(Glossary, layer = "data", assay = "RNA")[overlap, ]))
adt_glossary   <- as.data.frame(t(GetAssayData(Glossary, layer = "data", assay = "ADT")))
rna_glossary$pRPS6 <- adt_glossary$pRPS6

model <- lm(pRPS6 ~ ., data = rna_glossary)

# --- 5. Predict on Test object ---
rna_test       <- as.data.frame(t(GetAssayData(Test, layer = "data", assay = "RNA")[overlap, ]))
Test$pRPS6_OLS <- predict(model, newdata = rna_test)

#### LASSO model using Glossary Data and Applying to "Test" dataset
### First load in "Glossary", "Test" and "GWPS_mTOR" datasets
library(glmnet)
GWPS_counts <- GetAssayData(GWPS_mTOR@assays$RNA)
GWPS_genes <- rownames(GWPS_counts)

#########
# Helper to extract and transpose assay data
get_assay_matrix <- function(obj, assay, layer = "data") {
  as.data.frame(t(as.data.frame(GetAssayData(obj, layer = layer, assay = assay))))
}

# RNA counts
counts_data <- get_assay_matrix(Glossary, "RNA")
counts_test  <- get_assay_matrix(Test, "RNA")

# Filter to valid genes shared across both objects and GWPS_genes
valid_genes  <- Reduce(intersect, list(GWPS_genes, colnames(counts_data), colnames(counts_test)))
counts_data  <- as.matrix(counts_data[, valid_genes, drop = FALSE])

# ADT data
adt_data     <- get_assay_matrix(Glossary, "ADT")
resp_pRPS6   <- adt_data$pRPS6

cv_lasso <- cv.glmnet(
  counts_data, resp_pRPS6, 
  alpha = 1,               # alpha = 1 for LASSO
  family = "gaussian",     # Use "gaussian" for linear regression
  type.measure = "class",  # Measure for classification error
  nfolds = 10              # Number of cross-flex4 folds
)

# Lambda with the minimum cross-flex4 error
best_lambda <- cv_lasso$lambda.min

final_model_LASSO <- glmnet(
  counts_data, resp_pRPS6, 
  alpha = 1, 
  family = "gaussian", 
  lambda = best_lambda
)

############### Applying Model
counts_test  <- as.matrix(counts_test[, valid_genes, drop = FALSE])
predicted <- predict(final_model_LASSO, counts_test, type = "response")
Test$pRPS6_LASSO <- predicted 

#### Applying Ridge model, using the same valid genes as above

cv_ridge <- cv.glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0,               # alpha = 0 for Ridge
  family = "gaussian",     # Use "gaussian" for linear regression
  type.measure = "class",  # Measure for classification error
  nfolds = 10              # Number of cross-flex4 folds
)

# Lambda with the minimum cross-flex4 error
best_lambda <- cv_ridge$lambda.min

final_model_ridge <- glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0, 
  family = "gaussian", 
  lambda = best_lambda
)

############### Applying Model
predicted_probs_ridge <- predict(final_model_ridge, counts_test, type = "response")
Test$pRPS6_Ridge <- predicted_probs_ridge 
```

This just makes a scatterplot on the modeling data compared to the measured data in the test dataset

```{r Fig4A, dpi=300, fig.width=10, fig.height=5, dev='png'}
df <- as.data.frame(FetchData(Test, vars = c("pRPS6_Ridge","pRPS6_LASSO","pRPS6_OLS","pRPS6_measured")))
# Create the scatter plot with density-based color
p1 <- ggplot(df, aes(x = pRPS6_measured, y = pRPS6_OLS)) +
  stat_binhex(bins = 100) +  # Use hexbin for density-based coloring
  scale_fill_viridis_c(option = "plasma") +  # Use the 'plasma' viridis palette
  theme_minimal()+ 
  ylim(0,6) +
  xlim(0,6) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme(panel.grid = element_blank()) +
  labs(x = "adt_pRPS6",
       y = "pRPS6_OLS")

model <- lm(df$pRPS6_measured ~ df$pRPS6_OLS, data = df)
summary(model)

p2 <- ggplot(df, aes(x = pRPS6_measured, y = pRPS6_Ridge)) +
  stat_binhex(bins = 100) +  # Use hexbin for density-based coloring
  scale_fill_viridis_c(option = "plasma") +  # Use the 'plasma' viridis palette
  theme_minimal() +
  ylim(0,6) +
  xlim(0,6) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme(panel.grid = element_blank()) +
  labs(x = "adt_pRPS6",
       y = "pRPS6_Ridge")

model <- lm(df$pRPS6_measured ~ df$pRPS6_Ridge, data = df)
summary(model)

plot(p1+p2)
```

For the supplement, do the same boxplot but with the LASSO values

```{r FigS4A, dpi=300, fig.width=10, fig.height=5, dev='png'}
# Create the scatter plot with density-based color
p1 <- ggplot(df, aes(x = pRPS6_measured, y = pRPS6_LASSO)) +
  stat_binhex(bins = 100) +  # Use hexbin for density-based coloring
  scale_fill_viridis_c(option = "plasma") +  # Use the 'plasma' viridis palette
  theme_minimal()+ 
  ylim(0,6) +
  xlim(0,6) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme(panel.grid = element_blank()) +
  labs(x = "adt_pRPS6",
       y = "pRPS6_ridge")

model <- lm(df$pRPS6_measured ~ df$pRPS6_LASSO, data = df)
summary(model)

plot(p1)
```

Here we takes the coefficients from the model and multiply them by their average log-fold counts to get a model contribution score, plotting it on a rank plot

```{r Fig4B, dpi=300, fig.width=10, fig.height=5, dev='png'}
ridge_coef <- coef(final_model_ridge)
ridge_coef_df <- data.frame(
  feature = rownames(ridge_coef),
  coefficient = as.numeric(ridge_coef)
)

counts_data <- get_assay_matrix(Glossary, "RNA")
avg_log_count <- colMeans(counts_data)
avg_log_counts_df <- data.frame(
  gene = names(avg_log_counts),
  avg_log_count = as.numeric(avg_log_counts)
)

gene_weighted_scores <- avg_log_counts_df %>%
  inner_join(
    ridge_coef_df,
    by = c("gene" = "feature")
  ) %>%
  mutate(weighted_score = avg_log_count * coefficient)

gene_weighted_scores <- gene_weighted_scores %>%
  mutate(rank = rank(-weighted_score, ties.method = "min"))

library(dplyr)
library(ggplot2)
library(ggrepel)

plot_df <- gene_weighted_scores %>%
  arrange(desc(weighted_score)) %>%
  mutate(rank = row_number())

labels_to_show <- plot_df %>%
  slice_head(n = 15) %>%
  bind_rows(plot_df %>% slice_tail(n = 15)) %>%
  pull(gene)

p <- ggplot(plot_df, aes(x = rank, y = weighted_score)) +
  geom_point(color = "black", size = 1.5, alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_text_repel(
    aes(label = ifelse(gene %in% labels_to_show, gene, NA)),
    color = "red",
    size = 4,
    max.overlaps = Inf,
    box.padding = 0.4,
    point.padding = 0.3,
    segment.color = "red"
  ) +
  labs(
    x = "Rank of Gene (by Model Contribution Score)",
    y = "Model Contribution Score"
  ) +
  theme_minimal(base_size = 16)

p
```

We take the bottom 100 genes for the model contribution rank and run gene ontology on them using clusterprofiler

```{r Fig4C, dpi=300, fig.width=10, fig.height=5, dev='png'}
#cluster profiler plot for bottom 100 genes using model cntribution score

library(clusterProfiler)
library(enrichplot)
library(ggplot2)
library(org.Hs.eg.db)
library(dplyr)

bottom_100 <- gene_weighted_scores %>%
  arrange(weighted_score) %>%
  slice_head(n = 100)

ego <- enrichGO(
  gene          = bottom_100$gene,
  OrgDb         = org.Hs.eg.db,
  keyType       = "SYMBOL",
  ont           = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff  = 0.05,
  qvalueCutoff  = 0.2,
  readable      = TRUE
)

dotplot(ego, showCategory = 10)
```

Here, we run the model many different times using the top and bottom 'x' genes from the OLS, seeing how the predictions improve as more genes are added. 

```{r Fig4D, dpi=300, fig.width=10, fig.height=5, dev='png'}
df <- as.data.frame(FetchData(Test, vars = c("pRPS6_measured", "pRPS6_ridge_10","pRPS6_ridge_30","pRPS6_ridge_50","pRPS6_ridge_100","pRPS6_ridge_500","pRPS6_ridge_1000","pRPS6_ridge_2000","pRPS6_Ridge")))
model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge_10, data = df)
summary(model)
model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge_30, data = df)
summary(model)
model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge_50, data = df)
summary(model)
model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge_100, data = df)
summary(model)
model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge_500, data = df)
summary(model)
model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge_1000, data = df)
summary(model)
model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge_2000, data = df)
summary(model)
model <- lm(df$pRPS6_measured ~ df$pRPS6_Ridge, data = df)
summary(model)

# Example dataframe structure
df <- data.frame(
  top_n = factor(c(10, 30, 50, 100, 500, 1000, 2000, "All"), 
                 levels = c(10, 30, 50, 100, 500, 1000, 2000, "All")),
  correlation = c(0.65, 0.705, 0.711, 0.718,0.742, 0.750, 0.752, 0.786)
)

p1 <- ggplot(df, aes(x = top_n, y = correlation, group = 1)) +
  geom_line() +
  geom_point(size = 3) +
  labs(
    x = "Top/Bottom Genes in Model",
    y = "Correlation"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black")
  )

p1
```

This is a ridge plot showing the differences in predicted pRPS6 between TSC1 and RPTOR, where we should expect almost maximal change

```{r Fig4E, dpi=300, fig.width=10, fig.height=5, dev='png'}
p1 <- RidgePlot(GWPS_mTOR,"pRPS6_linear",idents = c("RPTOR","non-targeting","TSC1")) + xlim(1.5,6) + NoLegend() + theme(axis.title.y = element_blank(),axis.text.y = element_blank())

p2 <- RidgePlot(GWPS_mTOR,"pRPS6_ridge",idents = c("RPTOR","non-targeting","TSC1")) + xlim(1.5,6) + NoLegend() + theme(axis.title.y = element_blank(),axis.text.y = element_blank())

GWPS_mTOR_meta <- GWPS_mTOR@meta.data
GWPS_mTOR_meta_RT <- GWPS_mTOR_meta[GWPS_mTOR_meta$gene == c("RPTOR","TSC1"), ]
GWPS_mTOR_meta_NT <- GWPS_mTOR_meta[GWPS_mTOR_meta$gene == c("non-targeting","TSC1"), ]
GWPS_mTOR_meta_RN <- GWPS_mTOR_meta[GWPS_mTOR_meta$gene == c("RPTOR","non-targeting"), ]


t.test(pRPS6_linear ~ gene, data = GWPS_mTOR_meta_RT)
t.test(pRPS6_ridge ~ gene, data = GWPS_mTOR_meta_RT)
t.test(pRPS6_linear ~ gene, data = GWPS_mTOR_meta_NT)
t.test(pRPS6_ridge ~ gene, data = GWPS_mTOR_meta_NT)
t.test(pRPS6_linear ~ gene, data = GWPS_mTOR_meta_RN)
t.test(pRPS6_ridge ~ gene, data = GWPS_mTOR_meta_RN)

plot(p1 + p2)
```

Same thing but for the LASSO values
```{r FigS4B, dpi=300, fig.width=10, fig.height=5, dev='png'}

p3 <- RidgePlot(GWPS_mTOR,"pRPS6_LASSO",idents = c("RPTOR","non-targeting","TSC1")) + xlim(1.5,6) + NoLegend() + theme(axis.title.y = element_blank(),axis.text.y = element_blank())

t.test(pRPS6_LASSO ~ gene, data = GWPS_mTOR_meta_RT)
t.test(pRPS6_LASSO ~ gene, data = GWPS_mTOR_meta_NT)
t.test(pRPS6_LASSO ~ gene, data = GWPS_mTOR_meta_RN)

plot(p3)
```

Boxplot showing predicted pRPS6 in GWPS in a subset of perturbations present in the Glossary

```{r Fig4F, dpi=300, fig.width=10, fig.height=5, dev='png'}
GWPS_gloss <- readRDS("/brahms/blairj/FLEX/FLEX8/GWPS_gloss.rds")
Glossary_pRPS6_md <- GWPS_gloss@meta.data     

result_pRPS6 <- Glossary_pRPS6_md %>%
  group_by(gene) %>%
  summarise(median_pRPS6 = median(pRPS6_Ridge))

result_pRPS6 <- result_pRPS6[order(result_pRPS6$median_pRPS6, decreasing = TRUE), ,drop=F]

levels_pRPS6 <- result_pRPS6$gene

# Reorder the factor levels for the x-axis
Glossary_pRPS6_md$gene <- factor(Glossary_pRPS6_md$gene, levels = c(levels_pRPS6))


# Get default ggplot2 palette for all levels
library(scales)

n_idents <- length(unique(Glossary_pRPS6_md$gene))

default_cols <- hue_pal()(n_idents)

# Name them after the factor levels
names(default_cols) <- levels(Glossary_pRPS6_md$gene)

# Override only NO-TARGET color
default_cols["non-targeting"] <- "grey"

p1 <- ggplot(Glossary_pRPS6_md, aes(x = gene, y = pRPS6_Ridge, fill = gene)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  ylim(2, 5.5) +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

plot(p1)
```


Boxplot showing predicted pRPS6 in GWPS in a subset of mTOR relevant perturbations

```{r Fig4F, dpi=300, fig.width=10, fig.height=5, dev='png'}
mTOR_pRPS6_md <- GWPS_mtor@meta.data     

result_pRPS6 <- mTOR_pRPS6_md %>%
  group_by(gene) %>%
  summarise(median_pRPS6 = median(pRPS6_Ridge))

result_pRPS6 <- result_pRPS6[order(result_pRPS6$median_pRPS6, decreasing = TRUE), ,drop=F]

levels_pRPS6 <- result_pRPS6$gene

# Reorder the factor levels for the x-axis
mTOR_pRPS6_md$gene <- factor(mTOR_pRPS6_md$gene, levels = c(levels_pRPS6))

# Get default ggplot2 palette for all levels
library(scales)

n_idents <- length(unique(mTOR_pRPS6_md$gene))

default_cols <- hue_pal()(n_idents)

# Name them after the factor levels
names(default_cols) <- levels(mTOR_pRPS6_md$gene)

# Override only NO-TARGET color
default_cols["non-targeting"] <- "grey"

p1 <- ggplot(mTOR_pRPS6_md, aes(x = gene, y = pRPS6_Ridge, fill = gene)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  ylim(2, 5.5) +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

plot(p1)
```

This just makes a scatter plot with the median predicted pRPS6 values in GWPS vs the median measured pRSP6 values in the glossary dataset 

```{r FigS4D, dpi=300, fig.width=10, fig.height=5, dev='png'}
GWPS_gloss_md <- GWPS_gloss@meta.data
gloss_md <- Glossary@meta.data

median_data_GWPS <- GWPS_gloss_md %>%
  group_by(gene) %>%
  summarise(pRPS6_ridge = median(pRPS6_ridge))

median_data_gloss <- gloss_md %>%
  group_by(GDO_comb) %>%
  summarise(pRPS6_measured = median(pRPS6_measured))

names(median_data_gloss)[names(median_data_gloss) == "GDO_comb"] <- "gene"

gloss_merge <- merge(median_data_GWPS, median_data_gloss, by = "gene", all = FALSE)

ggplot(gloss_merge, aes(x = pRPS6_ridge, y = pRPS6_measured)) +
  # Add thicker zero lines
  geom_point(color = "black", size = 3) +# Thicker vertical line# Plot median points
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  theme_minimal()

model <- lm(gloss_merge$pRPS6_measured ~ gloss_merge$pRPS6_ridge, data = gloss_merge)
r2_value <- summary(model)$r.squared
summary(model)

library(ggrepel)

to_label <- c("TSC1","TSC2","ERCC5","RPTOR","DHDDS","GFI1B","TTI2","SEM1","VPS35","TRPM7","NPRL3")   # replace with your own

p1 <- ggplot(gloss_merge, aes(x = pRPS6_ridge, y = pRPS6_measured)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  geom_text_repel(
    data = subset(gloss_merge, gene %in% to_label),  # subset by vector
    aes(label = gene),
    color = "red",
    size = 4
  ) +
  theme_minimal()

plot(p1)
```

For this, you must downlaod the data from Opzoomer, et al., 2024, convert it into a seurat object with the measured pRPS6 values and run the ridge prediction model on it using the same subset of genes as above.  

```{r Fig4G, dpi=300, fig.width=10, fig.height=5, dev='png'}
SIGNAL_Hela <- readRDS("SIGNAL_Hela_pred.rds")

df <- SIGNAL_Hela@meta.data

FeatureScatter(SIGNAL_Hela,"adt_pRPS6","pRPS6_ridge",group.by = "orig.ident", cols = "black")+
  geom_smooth(method = "lm", color = "blue", se = TRUE) + NoLegend()

model <- lm(df$pRPS6_measured ~ df$pRPS6_ridge, data = df)
r2_value <- summary(model)$r.squared
summary(model)
```

For this, you must downlaod the data from Li et al, 2025, and run the prediction model using the same GWPS genes as above

```{r Fig4H, dpi=300, fig.width=10, fig.height=5, dev='png'}
# Load prediction object
obj_TL_pred <- readRDS("/brahms/blairj/references/obj_TL_pred.rds")

# Seurat matrices are genes x cells
mTOR_rna <- GetAssayData(obj_TL_pred, assay = "RNA", layer = "data")
gloss_rna <- GetAssayData(Glossary, assay = "RNA", layer = "data")

# Genes to use in model/prediction
common_genes <- Reduce(intersect, list(
  rownames(mTOR_rna),
  rownames(gloss_rna),
  GWPS_genes
))

# Keep genes in same order in both matrices
mTOR_rna <- mTOR_rna[common_genes, , drop = FALSE]
gloss_rna <- gloss_rna[common_genes, , drop = FALSE]

# Transpose once: cells x genes for glmnet
x_train <- t(gloss_rna)
x_pred  <- t(mTOR_rna)


adt_data <- GetAssayData(Glossary, assay = "ADT", layer = "data")


# Match response to training cells
resp_pRPS6 <- as.numeric(adt_data["pRPS6", colnames(gloss_rna)])

cv_ridge <- cv.glmnet(
  x = x_train,
  y = resp_pRPS6,
  alpha = 0,
  family = "gaussian",
  type.measure = "mse",
  nfolds = 10
)

best_lambda <- cv_ridge$lambda.min
lambda_1se  <- cv_ridge$lambda.1se

final_model_ridge <- glmnet(
  x = x_train,
  y = resp_pRPS6,
  alpha = 0,
  family = "gaussian",
  lambda = best_lambda
)

model_genes <- colnames(x_train)

# Check whether prediction matrix has all model genes
missing_genes <- setdiff(model_genes, colnames(x_pred))
extra_genes   <- setdiff(colnames(x_pred), model_genes)

length(missing_genes)
length(extra_genes)

# Reorder/filter prediction matrix to match training matrix
x_pred <- x_pred[, model_genes, drop = FALSE]

predicted_pRPS6 <- predict(
  final_model_ridge,
  newx = x_pred,
  type = "response"
)


predicted_pRPS6 <- as.numeric(predicted_pRPS6)
names(predicted_pRPS6) <- rownames(x_pred)

# Optional: add predictions to metadata
obj_TL_pred$pRPS6_ridge_pred <- predicted_pRPS6[colnames(obj_TL_pred)]

Idents(obj_TL_pred) <- "geno_time"

###boxplots
ObjTL_pRPS6_md <- obj_TL_pred@meta.data     
ObjTL_pRPS6_md$ident <- ObjTL_pRPS6_md$geno_time

# Reorder the factor levels for the x-axis
ObjTL_pRPS6_md$ident <- factor(ObjTL_pRPS6_md$ident, levels = c("CTL_50","TSC2KO_50","CTL_120","TSC2KO_120","CTL_220","TSC2KO_220"))



p1 <- ggplot(ObjTL_pRPS6_md, aes(x = ident, y = pRPS6_ridge_pred, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("CTL_50" = "grey", 
                               "TSC2KO_50" = "red3", 
                               "CTL_120"= "grey", 
                               "TSC2KO_120" = "red3", 
                               "CTL_220"= "grey",
                               "TSC2KO_220" = "red3")) +
  theme_minimal() + NoLegend() + theme(axis.title = element_blank()) + ylim(1.5,4)

plot(p1)

pairwise_results <- pairwise.t.test(
  x = ObjTL_pRPS6_md$pRPS6_ridge,
  g = ObjTL_pRPS6_md$ident,
  p.adjust.method = "BH",
  pool.sd = FALSE
)

pairwise_results
```

The following runs a 3' to Flex correction on the GWPS data to  account for any potential confounding effects of using comparing expression from the 10x Flex kit to the 10x 3' kit

```{r FigS4E, dpi=300, fig.width=10, fig.height=5, dev='png'}
######### Glossary instead
nt_GWPS <- subset(GWPS_mTOR,idents = "non-targeting")
nt_FLEX <- subset(Glossary,idents = "NO-TARGET")

gwps_nt <- GetAssayData(nt_GWPS,assay = "RNA",layer = "data")
flex_nt <- GetAssayData(nt_FLEX,assay = "RNA",layer = "data")
gwps_all <- GetAssayData(GWPS_mTOR,assay = "RNA",layer = "data")


common.genes <- Reduce(intersect, list(
  rownames(flex_nt),
  rownames(gwps_nt),
  rownames(gwps_all)
))

flex_nt <- flex_nt[common.genes, ]
gwps_nt <- gwps_nt[common.genes, ]
gwps_all <- gwps_all[common.genes, ]

mu_flex <- Matrix::rowMeans(flex_nt)
mu_gwps <- Matrix::rowMeans(gwps_nt)

delta <- mu_flex - mu_gwps

gwps_all_corrected <- sweep(
  gwps_all,
  MARGIN = 1,
  STATS = delta,
  FUN = "+"
)

row_sds <- function(x) {
  sqrt(Matrix::rowMeans(x^2) - Matrix::rowMeans(x)^2)
}

mu_flex <- Matrix::rowMeans(flex_nt)
mu_gwps <- Matrix::rowMeans(gwps_nt)

sd_flex <- row_sds(flex_nt)
sd_gwps <- row_sds(gwps_nt)

# avoid unstable ratios for low-variance genes
ratio <- sd_flex / sd_gwps
ratio[!is.finite(ratio)] <- 1
ratio[sd_gwps < 0.05] <- 1
ratio <- pmin(pmax(ratio, 0.25), 4)  # optional shrinkage/clipping

gwps_all_corrected <- sweep(gwps_all, 1, mu_gwps, "-")
gwps_all_corrected <- sweep(gwps_all_corrected, 1, ratio, "*")
gwps_all_corrected <- sweep(gwps_all_corrected, 1, mu_flex, "+")


#####
library(glmnet)
GWPS_counts <- GetAssayData(GWPS_mTOR@assays$RNA)
GWPS_genes <- rownames(GWPS_counts)

#########
# Helper to extract and transpose assay data
get_assay_matrix <- function(obj, assay, layer = "data") {
  as.data.frame(t(as.data.frame(GetAssayData(obj, layer = layer, assay = assay))))
}

# RNA counts
counts_data <- get_assay_matrix(Glossary, "RNA")

# Filter to valid genes shared across both objects and GWPS_genes
valid_genes  <- Reduce(intersect, list(common.genes, colnames(counts_data)))
counts_data  <- as.matrix(counts_data[, valid_genes, drop = FALSE])

# ADT data
adt_data     <- get_assay_matrix(Glossary, "ADT")
resp_pRPS6   <- adt_data$pRPS6


#### Applying Ridge model, using the same valid genes as above

cv_ridge <- cv.glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0,               # alpha = 1 for LASSO
  family = "gaussian",     # Use "gaussian" for linear regression
  type.measure = "class",  # Measure for classification error
  nfolds = 10              # Number of cross-flex4 folds
)

# Lambda with the minimum cross-flex4 error
best_lambda <- cv_ridge$lambda.min

final_model_ridge <- glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0, 
  family = "gaussian", 
  lambda = best_lambda
)

############### Applying Model
gwps_all_corrected <- t(gwps_all_corrected)
predicted_probs_ridge <- predict(final_model_ridge, gwps_all_corrected, type = "response")
GWPS_mTOR$pRPS6_Ridge_corrected <- predicted_probs_ridge 

mTOR_meta <- GWPS_mTOR@meta.data

median_data_mTOR <- mTOR_meta %>%
  group_by(gene) %>%
  summarise(
    med_pRPS6 = median(pRPS6_ridge),
    med_pRPS6_corrected = median(pRPS6_Ridge_corrected)
  )

#############
p2 <- ggplot(median_data_mTOR, aes(x = med_pRPS6, y = med_pRPS6_corrected)) +
  # Add thicker zero lines
  geom_point(color = "black", size = 3) +
  labs(
    title = "Predicted vs.Corrected",
    x = "Predicted pRPS6",
    y = "Corrected pRPS6"
  ) +
  theme_minimal()

model <- lm(median_data_mTOR$med_pRPS6 ~ median_data_mTOR$med_pRPS6_corrected, data = median_data_mTOR)
r2_value <- summary(model)$r.squared
R = sqrt(r2_value)
R = 0.96

plo(p2)
```

The following analyses runs and compares models generated with SCTransform-corrected data

```{r FigS4F, dpi=300, fig.width=10, fig.height=5, dev='png'}
#SCtransform modeling
library(glmnet)
GWPS_counts <- GetAssayData(GWPS_mTOR@assays$RNA)
GWPS_genes <- rownames(GWPS_counts)

##############
Glossary <- SCTransform(Glossary, return.only.var.genes = FALSE, method = "glmGamPoi")

#########
counts_data <- GetAssayData(Glossary,layer = "data",assay = "SCT")
counts_data  <- as.data.frame(t(as.data.frame(counts_data)))
valid_genes <- intersect(GWPS_genes, colnames(counts_data))
counts_data <- counts_data[, valid_genes, drop = FALSE]
counts_data <- as.matrix(counts_data)

adt_data <- GetAssayData(Glossary,layer = "data",assay = "ADT")
adt_data <- as.data.frame(t(as.data.frame(adt_data)))
resp_pRPS6 <- adt_data$pRPS6

cv_ridge <- cv.glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0,               # alpha = 1 for LASSO
  family = "gaussian",     # Use "gaussian" for linear regression
  type.measure = "class",  # Measure for classification error
  nfolds = 10              # Number of cross-flex4 folds
)

# Lambda with the minimum cross-flex4 error
best_lambda <- cv_ridge$lambda.min

final_model_ridge_pRPS6_SCT <- glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0, 
  family = "gaussian", 
  lambda = best_lambda
)

############### Applying Model
Test <- SCTransform(Test, return.only.var.genes = FALSE, method = "glmGamPoi")

counts <- GetAssayData(Test,layer = "data",assay = "SCT")
counts  <- as.data.frame(t(as.data.frame(counts)))
valid_genes <- intersect(valid_genes, colnames(counts))
counts <- counts[, valid_genes, drop = FALSE]
counts <- as.matrix(counts)

predicted_probs_ridge <- predict(final_model_ridge_pRPS6_SCT, counts, type = "response")
Test$pRPS6_SCT <- predicted_probs_ridge 

Test_adt <- as.data.frame(t(GetAssayData(Test,assay = "ADT")))
Test$pRPS6_measured <- Test_adt$pRPS6 

df <- as.data.frame(FetchData(Test, vars = c("pRPS6_SCT", "pRPS6_ridge","pRPS6_measured")))
# Create the scatter plot with density-based color
p1 <- ggplot(df, aes(x = pRPS6_measured, y = pRPS6_SCT)) +
  stat_binhex(bins = 100) +  # Use hexbin for density-based coloring
  scale_fill_viridis_c(option = "plasma") +  # Use the 'plasma' viridis palette
  theme_minimal()+ 
  ylim(0,6) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme(panel.grid = element_blank()) +
  labs(x = "adt_pRPS6",
       y = "pRPS6_SCT") + ggtitle("SCT vs Measured")

model <- lm(df$pRPS6_measured ~ df$Ridge_F7_SCT, data = df)
summary(model)

plot(p1)
```

```{r FigS4F, dpi=300, fig.width=10, fig.height=5, dev='png'}
df <- as.data.frame(FetchData(Test, vars = c("pRPS6_SCT", "pRPS6_ridge","pRPS6_measured")))
# Create the scatter plot with density-based color
p1 <- ggplot(df, aes(x = pRPS6_ridge, y = pRPS6_SCT)) +
  stat_binhex(bins = 100) +  # Use hexbin for density-based coloring
  scale_fill_viridis_c(option = "plasma") +  # Use the 'plasma' viridis palette
  theme_minimal()+ 
  ylim(0,6) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme(panel.grid = element_blank()) +
  labs(x = "adt_pRPS6",
       y = "pRPS6_SCT") + ggtitle("SCT vs Measured")

model <- lm(df$pRPS6_measured ~ df$Ridge_F7_SCT, data = df)
summary(model)

plot(p1)
```

```{r FigS4H, dpi=300, fig.width=10, fig.height=5, dev='png'}
#SCtransform modeling
library(glmnet)
GWPS_counts <- GetAssayData(GWPS_mTOR@assays$RNA)
GWPS_genes <- rownames(GWPS_counts)

##############
Glossary <- SCTransform(Glossary, return.only.var.genes = FALSE, method = "glmGamPoi")

#########
counts_data <- GetAssayData(Glossary,layer = "data",assay = "SCT")
counts_data  <- as.data.frame(t(as.data.frame(counts_data)))
valid_genes <- intersect(GWPS_genes, colnames(counts_data))
counts_data <- counts_data[, valid_genes, drop = FALSE]
counts_data <- as.matrix(counts_data)

adt_data <- GetAssayData(Glossary,layer = "data",assay = "ADT")
adt_data <- as.data.frame(t(as.data.frame(adt_data)))
resp_pRPS6 <- adt_data$pRPS6

cv_ridge <- cv.glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0,               # alpha = 1 for LASSO
  family = "gaussian",     # Use "gaussian" for linear regression
  type.measure = "class",  # Measure for classification error
  nfolds = 10              # Number of cross-validation folds
)

# Lambda with the minimum cross-validation error
best_lambda <- cv_ridge$lambda.min

final_model_ridge_pRPS6_SCT <- glmnet(
  counts_data, resp_pRPS6, 
  alpha = 0, 
  family = "gaussian", 
  lambda = best_lambda
)

############### Applying Model
GWPS_mTOR <- JoinLayers(GWPS_mTOR)
GWPS_mTOR <- SCTransform(GWPS_mTOR, return.only.var.genes = FALSE, method = "glmGamPoi")

counts <- GetAssayData(GWPS_mTOR,layer = "data",assay = "SCT")
counts  <- as.data.frame(t(as.data.frame(counts)))
valid_genes <- intersect(valid_genes, colnames(counts))
counts <- counts[, valid_genes, drop = FALSE]
counts <- as.matrix(counts)

predicted_probs_ridge <- predict(final_model_ridge_pRPS6_SCT, counts, type = "response")
GWPS_mTOR$pRPS6_SCT <- predicted_probs_ridge

mTOR_pRPS6_md <- GWPS_mtor@meta.data     

result_pRPS6 <- mTOR_pRPS6_md %>%
  group_by(gene) %>%
  summarise(median_pRPS6 = median(pRPS6_SCT))

result_pRPS6 <- result_pRPS6[order(result_pRPS6$median_pRPS6, decreasing = TRUE), ,drop=F]

levels_pRPS6 <- result_pRPS6$gene

# Reorder the factor levels for the x-axis
mTOR_pRPS6_md$gene <- factor(mTOR_pRPS6_md$gene, levels = c(levels_pRPS6))

# Get default ggplot2 palette for all levels
library(scales)

n_idents <- length(unique(mTOR_pRPS6_md$gene))

default_cols <- hue_pal()(n_idents)

# Name them after the factor levels
names(default_cols) <- levels(mTOR_pRPS6_md$gene)

# Override only NO-TARGET color
default_cols["non-targeting"] <- "grey"

p1 <- ggplot(mTOR_pRPS6_md, aes(x = gene, y = pRPS6_SCT, fill = gene)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  ylim(2, 5.5) +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

plot(p1)
```

```{r FigS4I, dpi=300, fig.width=10, fig.height=5, dev='png'}
# Extract the metadata from the Seurat object
GWPS_meta <- GWPS_mTOR@meta.data

# Calculate medians for each gene
median_data_GWPS <- GWPS_meta %>%
  group_by(gene) %>%
  summarise(
    median_pRPS6_ridge = median(pRPS6_Ridge),
    median_pRPS6_SCT = median(pRPS6_SCT)
    
  )

idents_df <- as.data.frame(table(Idents(GWPS_mTOR)))


###########
labels_to_show <- c("TSC1","DEPDC5","ITFG2","RHEB","RPTOR","MTOR","MED12","MED19","NPRL2",
                    "NPRL3","EFI2B5","TTI1","TELO2","LAMTOR3","LAMTOR2","XRN1","SMC3","GFI1B","KDM1A","VIM")

# Plot only the medians with repelling labels
p1 <- ggplot(median_data_GWPS, aes(y = median_pRPS6_ridge, x = median_pRPS6_SCT)) +
  geom_point(color = "black", size = 3) +# Thicker vertical line# Plot median points
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  geom_text_repel(aes(label = ifelse(gene %in% labels_to_show, gene, NA)),
                  color = "red", size = 4,
                  box.padding = 0.5, max.overlaps = Inf
  ) +
  labs(
    title = "Predicted vs. Measured - Log",
    x = "Predicted pRPS6 - Log",
    y = "Predicted pRPS6 - SCT"
  ) +
  theme_minimal()

plot(p1)

# Fit a linear model
model <- lm(median_data_GWPS$median_pRPS6_SCT ~ median_data_GWPS$median_pRPS6_ridge, data = median_data_GWPS)

# Extract R² value
r2_value <- summary(model)$r.squared
r2 = 0.88
```
