---
title: "Fig5"
output: html_document
date: "2026-06-11"
---
---
title: "Figure 5 plots"
output: github_document
date: "2026-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(tidyr)
library(dplyr)
library(Matrix)
library(stringr)
library(ggrepel)
library(readr)
library(cowplot)
library(ggplot2)
library(tibble)
library(scales)



options(future.globals.maxSize = 10 * 1024^3)
```


``` {r Fig5B, dpi=300, fig.width=10, fig.height=5, dev='png'}
GWPS_md <- read_csv("/brahms/blairj/FLEX/FLEX8/GWPS_metadata.csv")

genes_keep <- c(
  "TSC1", "ITFG2", "DEPDC5", "FOXN1", "TELO2", "ATP6V1A",
  "non-targeting", "NARS", "SIRT7", "MTOR", "SRP72", "MED12",
  "MED19", "EIF2B5", "ASCC3", "DAD1", "EEF1G", "GPN3", "TAF1",
  "TAF8", "KDM1A", "SMC3", "EIF3D", "SUPT6H", "AARS", "RPTOR",
  "DHDDS"
)

GWPS_md_box <- subset(GWPS_md, gene %in% genes_keep)

median_data_gwps <- GWPS_md_box %>%
  group_by(gene) %>%
  summarise(
    median_pRPS6 = median(pRPS6_ridge))

median_data_gwps <- median_data_gwps[order(median_data_gwps$median_pRPS6, decreasing = TRUE), ,drop=F]

levels_pRPS6 <- median_data_gwps$gene

# Reorder the factor levels for the x-axis
GWPS_md_box$gene <- factor(GWPS_md_box$gene, levels = c(levels_pRPS6))

# Get default ggplot2 palette for all levels
default_cols <- hue_pal()(length(genes_keep))

# Name them after the factor levels
names(default_cols) <- levels(GWPS_md_box$gene)

# Override only NO-TARGET color
default_cols["non-targeting"] <- "grey"

p1 <- ggplot(GWPS_md_box, aes(x = gene, y = pRPS6_ridge, fill = gene)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  ylim(1.5, 5) +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

plot(p1)
```


```{r Fig5C, dpi=300, fig.width=10, fig.height=5, dev='png'}
# Example data
genes <- TvB.gene_summary$id
ranks <- TvB.gene_summary$neg.rank # Example ranks
genes_of_interest <- low_genes_pRPS6  # Genes to highlight

# Create a data frame
data <- data.frame(
  gene = genes,
  rank = ranks,
  highlight = ifelse(genes %in% genes_of_interest, "red", "gray")  # Highlight color
)

# Divide the data into 10 bins based on rank
data2 <- data %>%
  mutate(
    bin = cut(rank, breaks = 50, labels = FALSE)  # Create bins
  )

# Filter data for only red genes
red_genes_data <- data2 %>% dplyr::filter(highlight == "red")

# Plot histogram for red genes
p2 <- ggplot(red_genes_data, aes(x = bin)) +
  geom_histogram(binwidth = 1, fill = "red", color = "black") +  # Histogram for red genes
  labs(
    title = "Histogram of Highlighted Genes by Rank Bin",
    x = "Rank Bin",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove all grid lines
    axis.line = element_line(color = "black")  # Optional: add axis lines for clarity
  )

plot(p2)
```

To cluster the genes for validation selection, we ran mixscape and LDA on the bottom 200 and top 50 perturbations

```{r FigS5A, dpi=300, fig.width=10, fig.height=5, dev='png'}
DefaultAssay(object = GWPS_ridge) <- 'RNA'
GWPS_ridge <- NormalizeData(object = GWPS_ridge) %>% FindVariableFeatures() %>% ScaleData()

# Run Principle Component Analysis (PCA) to reduce the dimensionality of the data.
GWPS_ridge <- RunPCA(object = GWPS_ridge)

# Run Uniform Manifold Approximation and Projection (UMAP) to visualize clustering in 2-D.
GWPS_ridge <- RunUMAP(object = GWPS_ridge, dims = 1:40)

Idents(GWPS_ridge) <- "GDO_comb"

# Calculate perturbation signature (PRTB).
GWPS_ridge<- CalcPerturbSig(
  object = GWPS_ridge, 
  assay = "RNA", 
  slot = "data", 
  gd.class ="gene", 
  nt.cell.class = "non-targeting", 
  reduction = "pca", 
  ndims = 40, 
  num.neighbors = 20, 
  new.assay.name = "PRTB")

# Prepare PRTB assay for dimensionality reduction: 
# Normalize data, find variable features and center data.
DefaultAssay(object = GWPS_ridge) <- 'PRTB'

# Use variable features from RNA assay.
VariableFeatures(object = GWPS_ridge) <- VariableFeatures(object = GWPS_ridge[["RNA"]])
GWPS_ridge <- ScaleData(object = GWPS_ridge, do.scale = F, do.center = T)

# Run PCA to reduce the dimensionality of the data.
GWPS_ridge <- RunPCA(object = GWPS_ridge, reduction.key = 'prtbpca', reduction.name = 'prtbpca')

# Run mixscape.
GWPS_ridge <- RunMixscape(
  object = GWPS_ridge, 
  assay = "PRTB", 
  slot = "scale.data", 
  labels = "gene", 
  nt.class.name = "non-targeting", 
  min.de.genes = 1, 
  iter.num = 10, 
  de.assay = "RNA", 
  verbose = F,
  prtb.type = "KO",
  logfc.threshold = 0.05)

# Run DE analysis and visualize results on a heatmap ordering cells by their posterior 
# probability values.

# Remove non-perturbed cells and run LDA to reduce the dimensionality of the data.
Idents(GWPS_ridge) <- "mixscape_class.global"
sub <- subset(GWPS_ridge, idents = c("KO", "NO"))

sub <- MixscapeLDA(
  object = sub, 
  assay = "RNA", 
  pc.assay = "PRTB", 
  labels = "gene", 
  nt.label = "non-targeting", 
  npcs = 20, 
  logfc.threshold = 0.05, 
  verbose = F)

# Use LDA results to run UMAP and visualize cells on 2-D. 
# Here, we note that the number of the dimensions to be used is equal to the number of 
# labels minus one (to account for NT cells).


sub <- RunUMAP(
  object = sub,
  dims = 1:40,
  reduction = 'lda',
  reduction.key = 'ldaprtbumap',
  reduction.name = 'ldaprtbumap')


Idents(sub) <- "GDO_comb"


p1 <- DimPlot(object = sub, 
              reduction = "ldaprtbumap",label = T,label.box = T,pt.size = 1,label.size = 3) + NoLegend() + theme(axis.title = element_blank(),axis.ticks = element_blank(),axis.text = element_blank())

plot(p1)

```

To ensure that our starvation paradigm was not causing poor predictions from the GWPS data that was not starved, we ran another screen, analyzed with mageck with feeding and starving the cells as described in the methods. 

```{r FigS5B, dpi=300, fig.width=10, fig.height=5, dev='png'}
Feed.gene_summary <- read.delim("/brahms/blairj/PhosphoScreen/StarveFeedValidPred/mageck/Feed.gene_summary.txt")
Starve.gene_summary <- read.delim("/brahms/blairj/PhosphoScreen/StarveFeedValidPred/mageck/Starve.gene_summary.txt")

merged_summary <- merge(Feed.gene_summary,Starve.gene_summary, by = "id", all = FALSE)

# Plot only the medians with repelling labels
p1 <- ggplot(merged_summary, aes(x = neg.lfc.x, y = neg.lfc.y)) +
  # Add thicker zero lines
  geom_point(color = "black", size = 3) +# Thicker vertical line# Plot median points
  labs(
    title = "Validation vs. Validation - Bulk Screen",
    x = "Feed Validation Pool Log2FC",
    y = "Starve Validation Pool Log2FC"
  ) +
  theme_minimal()

model <- lm(merged_summary$neg.lfc.x ~ merged_summary$neg.lfc.y, data = merged_summary)
r2_value <- summary(model)$r.squared
R = sqrt(r2_value)
R = 0.95
```


```{r Fig5D, dpi=300, fig.width=10, fig.height=5, dev='png'}
Idents(Validation) <- "GDO_comb"
Validation_box <- subset(Validation,idents = c("TSC1","ITFG2","DEPDC5","FOXN1","TELO2","ATP6V1A","NO","NARS","SIRT7","MTOR","SRP72","MED12","MED19","EIF2B5","ASCC3","DAD1","EEF1G","GPN3","TAF1","TAF8","KDM1A","SMC3","EIF3D","SUPT6H","AARS","RPTOR","DHDDS"))

###boxplots sane thing
Validation_box_pRPS6 <- GetAssayData(Validation_box,assay = "ADT")
Validation_box_pRPS6 <- as.data.frame(Validation_box_pRPS6)
Validation_box_pRPS6 <- as.data.frame(t(Validation_box_pRPS6))
Validation_box_pRPS6_md <- Validation_box@meta.data     
Validation_box_pRPS6$ident <- Validation_box_pRPS6_md$GDO_comb

result_pRPS6 <- Validation_box_pRPS6_md %>%
  group_by(GDO_comb) %>%
  summarise(median_pRPS6 = median(pRPS6_measured))

result_pRPS6 <- result_pRPS6[order(result_pRPS6$median_pRPS6, decreasing = TRUE), ,drop=F]

levels_pRPS6 <- result_pRPS6$GDO_comb

# Reorder the factor levels for the x-axis
Validation_box_pRPS6$ident <- factor(Validation_box_pRPS6$ident, levels = c("DEPDC5","TSC1","FOXN1","ITFG2","NO","NARS","EIF3D","ASCC3","TAF8","ATP6V1A","GPN3","DAD1","EEF1G","RPTOR","MTOR","SRP72","TAF1","EIF2B5","TELO2","KDM1A","SIRT7","DHDDS","SMC3","MED19","AARS","SUPT6H","MED12"))

# Get default ggplot2 palette for all levels
default_cols <- hue_pal()(length(unique(Validation_box_pRPS6$ident)))

# Name them after the factor levels
names(default_cols) <- levels(Validation_box_pRPS6$ident)

# Override only NO-TARGET color
default_cols["NO"] <- "grey"

p2 <- ggplot(Validation_box_pRPS6, aes(x = ident, y = pRPS6, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  ylim(2, 5.5) +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())
plot(p2)

```


```{r Fig5E, dpi=300, fig.width=10, fig.height=5, dev='png'}
Idents(Validation) <- "GDO_classification.global"
GDO_class <- as.data.frame(table(Idents(Validation)))

GDO_f8 <- subset(Validation, idents = c("NO-TARGET-177","NO-TARGET-108","SMC3-2","SMC3-4","TAF8-3","TAF8-0","MED12-0","MED12-5"))
levels(GDO_f8) <- c("NO-TARGET-177","NO-TARGET-108","SMC3-2","SMC3-4","TAF8-3","TAF8-0","MED12-0","MED12-5")


###boxplots
GDO_f8_pRPS6 <- GetAssayData(GDO_f8,assay = "ADT")
GDO_f8_pRPS6 <- as.data.frame(GDO_f8_pRPS6)
GDO_f8_pRPS6 <- as.data.frame(t(GDO_f8_pRPS6))
                    
GDOf8_md <- GDO_f8@meta.data     
GDO_f8_pRPS6$ident <- GDOf8_md$GDO_classification.global

library(tidyverse)
library(ggplot2)

# Reorder the factor levels for the x-axis
GDO_f8_pRPS6$ident <- factor(GDO_f8_pRPS6$ident, levels = c("NO-TARGET-177","NO-TARGET-108","SMC3-2","SMC3-4","TAF8-3","TAF8-0","MED12-0","MED12-5"))

# Create boxplot with custom fill colors
p1 <- ggplot(GDO_f8_pRPS6, aes(x = ident, y = pRPS6, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("NO-TARGET-13" = "grey55", 
                               "NO-TARGET-125" = "grey55", 
                               "SMC3-2"= "#00BA38","SMC3-4" = "#00BA38", 
                               "TAF8-3" = "#619CFF", 
                               "TAF8-0"= "#619CFF",
                               "MED12-0" = "#F8766D", 
                               "MED12-5"= "#F8766D")) +
  theme_minimal() + NoLegend()

png("plots/BoxPlot_byGuideValid.png",width = 2400, height = 1500,res=300)
plot(p1)
dev.off()

##significance calculation
control_group <- "NO-TARGET-177"

pairwise_results <- pairwise.t.test(
  x = GDO_f8_pRPS6$pRPS6,
  g = GDO_f8_pRPS6$ident,
  p.adjust.method = "BH",
  pool.sd = FALSE
)

pairwise_results

control_vals <- GDO_f8_pRPS6$pRPS6[GDO_f8_pRPS6$ident == "NO-TARGET-177"]
mean_control <- mean(control_vals)
results <- GDO_f8_pRPS6 %>%
  filter(ident != "NO-TARGET-177") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(pRPS6),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(pRPS6, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results

```


``` {r Fig5F, dpi=300, fig.width=10, fig.height=5, dev='png'}
########
G_meta <- GWPS_flex8genes@meta.data

# Calculate medians for each gene
median_data_GWPS <- G_meta %>%
  group_by(gene) %>%
  summarise(median_pRPS6 = median(pRPS6_ridge))

adt_data <- as.data.frame(t(as.data.frame(GetAssayData(FLEX8,assay = "ADT"))))
FLEX8$pRPS6_measured <- adt_data$pRPS6

# Extract the metadata from the Seurat object
FLEX_meta <- FLEX8@meta.data

# Calculate medians for each gene
median_data_FLEX <- FLEX_meta %>%
  group_by(GDO_comb) %>%
  summarise(
    median_pRPS6_measured = median(pRPS6_measured),
    median_pRPS6_FLEX7 = median(pRPS6_ridge_GWPS),
  )

rownames(median_data_GWPS) <- median_data_GWPS$gene
nt_med <- median_data_GWPS["non-targeting","median_pRPS6"]
nt <- nt_med[[1]][1]
median_data_GWPS$delta_pRPS6 <- median_data_GWPS$median_pRPS6 - nt

rownames(median_data_FLEX) <- median_data_FLEX$GDO_comb
nt_med <- median_data_FLEX["NO","median_pRPS6_measured"]
nt <- nt_med[[1]][1]
median_data_FLEX$delta_pRPS6 <- median_data_FLEX$median_pRPS6_measured - nt

rownames(median_data_FLEX) <- median_data_FLEX$GDO_comb
nt_med <- median_data_FLEX["NO","median_pRPS6_FLEX7"]
nt <- nt_med[[1]][1]
median_data_FLEX$delta_pRPS6_FLEX7 <- median_data_FLEX$median_pRPS6_FLEX7 - nt

idents_df <- as.data.frame(table(Idents(FLEX8)))

merged_df1 <- median_data_GWPS[, c("gene","delta_pRPS6","median_pRPS6")]
colnames(merged_df1) <- c("Var1","delta_pRPS6_GWPS","median_pRPS6_GWPS")

merged_df2 <- median_data_FLEX[, c("GDO_comb","delta_pRPS6","delta_pRPS6_FLEX7","median_pRPS6_measured","median_pRPS6_FLEX7")]
colnames(merged_df2) <- c("Var1","delta_pRPS6_meas","delta_pRPS6_FLEX7","median_pRPS6_meas","median_pRPS6_FLEX7")

merged_pool_3 <- merge(merged_df1,merged_df2, by = "Var1", all = FALSE)

labels_show <- idents_df$Var1

labels_to_show1 <- c("TSC1","DEPDC5","ITFG2","NO","RHEB","RPTOR","MTOR","MED12","MED19","NPRL2",
                     "NPRL3","EFI2B5","TTI1","TELO2","LAMTOR3","LAMTOR2","XRN1","TAF8","SMC3","GFI1B","KDM1A")
# Plot only the medians with repelling labels
p8 <- ggplot(merged_pool_3, aes(x = median_pRPS6_FLEX7, y = median_pRPS6_meas)) +
  # Add thicker zero lines
  geom_hline(yintercept = 0, color = "grey", size = 1.2) +  # Thicker horizontal line
  geom_vline(xintercept = 0, color = "grey", size = 1.2) + 
  geom_point(color = "black", size = 3) +# Thicker vertical line# Plot median points
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  geom_text_repel(aes(label = ifelse(Var1 %in% labels_to_show1, Var1, NA)),
                  color = "red", size = 4,
                  box.padding = 0.5, max.overlaps = Inf
  ) +
  ylim(2.5,5) + xlim(2.5,4.5) +
  labs(
    title = "8 - Predicted vs. Measured - FLEX7 v FLEX8 - No MS -GWPS only - median",
    x = "Delta Predicted pRPS6 FLEX7",
    y = "Delta Measured pRPS6"
  ) +
  theme_minimal()

plot(p8)
```

The following plots aare all based on the modeling of VIM, which uses the same methods as the modeling of pRPS6

``` {r FigS5C, dpi=300, fig.width=10, fig.height=5, dev='png'}
counts_data <- GetAssayData(Glossary,layer = "data",assay = "RNA")
counts_data  <- as.data.frame(t(as.data.frame(counts_data)))
counts_data <- subset(counts_data[,GWPS_genes])
counts_data <- as.matrix(counts_data)

adt_data <- GetAssayData(Glossary,layer = "data",assay = "ADT")
adt_data <- as.data.frame(t(as.data.frame(adt_data)))
resp_VIM <- adt_data$VIM

cv_ridge <- cv.glmnet(
  counts_data, resp_VIM, 
  alpha = 0,               # alpha = 1 for LASSO
  family = "gaussian",     # Use "gaussian" for linear regression
  type.measure = "class",  # Measure for classification error
  nfolds = 10              # Number of cross-validation folds
)

# Lambda with the minimum cross-validation error
best_lambda <- cv_ridge$lambda.min

final_model_ridge <- glmnet(
  counts_data, resp_VIM, 
  alpha = 0, 
  family = "gaussian", 
  lambda = best_lambda
)

counts <- GetAssayData(Test,layer = "data",assay = "RNA")
counts <- t(counts)
counts <- subset(counts[,GWPS_genes])
counts <- as.matrix(counts)

predicted_probs_ridge <- predict(final_model_VIM, counts, type = "response")
Test$VIM_ridge <- predicted_probs_ridge 

adt_data <- GetAssayData(flex4,layer = "data",assay = "ADT")
adt_data <- as.data.frame(t(as.data.frame(adt_data)))
Test$VIM_measured <- adt_data$VIM

#######
counts <- GetAssayData(Validation,layer = "data",assay = "RNA")
counts <- t(counts)
counts <- subset(counts[,GWPS_genes])
counts <- as.matrix(counts)

predicted_probs_ridge <- predict(final_model_VIM, counts, type = "response")
Validation$VIM_ridge <- predicted_probs_ridge 

df <- as.data.frame(FetchData(Test, vars = c("VIM_measured","VIM_ridge")))
# Create the scatter plot with density-based color
p1 <- ggplot(df, aes(x = VIM_measured, y = VIM_ridge)) +
  stat_binhex(bins = 100) +  # Use hexbin for density-based coloring
  scale_fill_viridis_c(option = "plasma") +  # Use the 'plasma' viridis palette
  theme_minimal()+ 
  ylim(0,6) +
  xlim(0,6) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  theme(panel.grid = element_blank()) +
  labs(x = "adt_pRPS6",
       y = "pRPS6_OLS")

model <- lm(df$pRPS6_measured ~ df$pRPS6_OLS, data = df)
summary(model)

plot(p1)

```

``` {r FigS5D, dpi=300, fig.width=10, fig.height=5, dev='png'}
Validation_md <- Validation@meta.data

median_data_valid <- Validation_md %>%
  group_by(GDO_comb) %>%
  summarise(VIM_measured = median(VIM_measured),VIM_ridge = median(VIM_ridge))

model <- lm(median_data_valid$VIM_measured ~ median_data_valid$VIM_ridge, data = median_data_valid)
r2_value <- summary(model)$r.squared
summary(model)

library(ggrepel)

to_label <- c("TSC1","TSC2","ERCC5","RPTOR","DHDDS","GFI1B","TTI2","SEM1","VPS35","TRPM7","NPRL3")   # replace with your own

p1 <- ggplot(median_data_valid, aes(x = VIM_ridge, y = VIM_measured)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  geom_text_repel(
    data = subset(gloss_merge, gene %in% to_label),  # subset by vector
    aes(label = gene),
    color = "red",
    size = 4
  ) +
  theme_minimal()

plot(p1)
```

``` {r FigS5E, dpi=300, fig.width=10, fig.height=5, dev='png'}
Validation_Vim <- GetAssayData(Validation,assay = "ADT")
Validation_VIM <- as.data.frame(Validation_VIM)
Validation_VIM <- as.data.frame(t(Validation_VIM))
Validation_VIM_md <- Validation@meta.data     
Validation_VIM$ident <- Validation_VIM_md$GDO_comb

result_VIM <- Validation_VIM_md %>%
  group_by(GDO_comb) %>%
  summarise(median_VIM = median(VIM_measured))

result_VIM <- result_VIM[order(result_VIM$median_VIM, decreasing = TRUE), ,drop=F]

levels_VIM <- result_VIM$GDO_comb

# Reorder the factor levels for the x-axis
Validation_VIM$ident <- factor(Validation_VIM$ident, levels = c(levels_VIM))

# Get default ggplot2 palette for all levels
library(scales)

n_idents <- length(unique(Validation_VIM$ident))

default_cols <- hue_pal()(n_idents)

# Name them after the factor levels
names(default_cols) <- levels(Validation_VIM$ident)

# Override only NO-TARGET color
default_cols["NO-TARGET"] <- "grey"

p1 <- ggplot(Validation_VIM, aes(x = ident, y = VIM, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  ylim(2, 5.5) +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

plot(p1)
```


``` {r FigS5F, dpi=300, fig.width=10, fig.height=5, dev='png'}
Idents(Validation) <- "GDO_comb"
vim <- AggregateExpression(Validation,assays = "RNA",return.seurat = T)
vim <- NormalizeData(vim)

avg_exp <- GetAssayData(vim,assay = "RNA")
avg_exp <- as.data.frame(avg_exp)


p1 <- ggplot(avg_exp, aes(x = NO, y = VIM)) +
  geom_point(color = "black", size = 3) +
  geom_text(aes(label = ifelse(rownames(avg_exp) == "CD55", "CD55", "")),
            hjust = -0.2, vjust = -0.5, color = "red") +
  geom_hline(yintercept = 0, color = "black") +
  geom_vline(xintercept = 0, color = "black") +
  theme_minimal() +
  theme(
    panel.grid = element_blank()  # Remove gridlines
  )

plot(p1)
```

The following uses the Vimentin bulk screen data

``` {r FigS5G, dpi=300, fig.width=10, fig.height=5, dev='png'}
TvB.gene_summary <- read.delim("Top.gene_summary.txt")
TvB.gene_summary$smaller_value <- pmin(TvB.gene_summary$neg.p.value, TvB.gene_summary$pos.p.value)
TvB.gene_summary$smaller_rank <- pmin(TvB.gene_summary$neg.rank, TvB.gene_summary$pos.rank)
TvB.gene_summary$smaller_rank <- (TvB.gene_summary$smaller_rank)/18939
genes_lab <- c("VIM","KDM1A","GFI1B","LAMTOR2","LAMTOR3")

p1 <- EnhancedVolcano(TvB.gene_summary,TvB.gene_summary$id,"neg.lfc","smaller_value",drawConnectors = T,arrowheads = F,selectLab = genes_lab,xlim = c(-6.5,6.5),ylim = c(0,7),FCcutoff = 1,pCutoff = 0.05,title = NULL,subtitle = NULL)

plot(p1)
```

``` {r FigS5H, dpi=300, fig.width=10, fig.height=5, dev='png'}
library(ggplot2)

median_df <- GWPS_metadata_predictions %>%
  group_by(gene) %>%
  summarize(
    median_VIM   = median(VIM_ridge,   na.rm = TRUE)
  )

median_gwps <- median_df[order(median_df$median_VIM, decreasing = TRUE), ,drop=F]
high_genes_VIM <- median_gwps[1:200,]
high_genes_VIM <- as.character(high_genes_VIM$gene)

# Example data
Top.gene_summary <- read.delim("/Top.gene_summary_vim.txt")
genes <- Top.gene_summary$id
ranks <- Top.gene_summary$pos.rank # Example ranks
genes_of_interest <- high_genes_VIM  # Genes to highlight

# Create a data frame
data <- data.frame(
  gene = genes,
  rank = ranks,
  highlight = ifelse(genes %in% genes_of_interest, "red", "gray")  # Highlight color
)

# Divide the data into 50 bins based on rank
data2 <- data %>%
  mutate(
    bin = cut(rank, breaks = 50, labels = FALSE)  # Create bins
  )

# Filter data for only red genes
red_genes_data <- data2 %>% dplyr::filter(highlight == "red")

# Plot histogram for red genes
p5 <- ggplot(red_genes_data, aes(x = bin)) +
  geom_histogram(binwidth = 1, fill = "red", color = "black") +  # Histogram for red genes
  labs(
    title = "Histogram of Highlighted Genes by Rank Bin",
    x = "Rank Bin",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove all grid lines
    axis.line = element_line(color = "black")  # Optional: add axis lines for clarity
  )

plot(p5)

####
table(red_genes_data$bin)

# Observed: 37 in bin 1, 163 in other bins
observed <- c(35, 165)

# Expected: uniform distribution would be 20 in bin 1, 180 in others
expected <- c(20, 180)

# Chi-squared test
chisq.test(x = observed, p = expected / sum(expected))
```