---
title: "Fig6"
output: html_document
date: "2026-06-11"
---
---
title: "Figure 6 plots"
output: github_document
date: "2026-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(tidyr)
library(dplyr)
library(Matrix)
library(stringr)
library(ggrepel)
library(readr)
library(cowplot)
library(ggplot2)
library(tibble)
library(scales)

options(future.globals.maxSize = 10 * 1024^3)
```


Here is where we use k-means clustering from a subset of genes on a subset of perturbations to illustrate the related biological processes. First Read in the validation dataset

``` {r Fig6A, dpi=300, fig.width=10, fig.height=5, dev='png'}
####### Finding the input genes
Idents(Validation) <- "GDO_comb"
MTOR_DE <- FindMarkers(Validation,ident.2 = "NO",ident.1 = c("MTOR","RPTOR"),only.pos = F)
MED_DE <- FindMarkers(Validation,ident.2 = "NO",ident.1 = c("MED12","MED19","MED30"),only.pos = F)
TSC1_DE <- FindMarkers(Validation,ident.2 = "NO",ident.1 = c("TSC1"),only.pos = F)

MTOR_DE %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  slice_head(n = 20) %>%
  ungroup() -> MTOR_DE
MTOR_DE$gene <- rownames(MTOR_DE)

MED_DE %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  slice_head(n = 20) %>%
  ungroup() -> MED_DE
MED_DE$gene <- rownames(MED_DE)

TSC1_DE %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  slice_head(n = 20) %>%
  ungroup() -> TSC1_DE
TSC1_DE$gene <- rownames(TSC1_DE)

ATF6_UPR <- read.table("files/ATF6_UPR.txt")
IRE1a_UPR <- read.table("files/IRE1a_UPR.txt")
PERK_up <- read.table("files/PERK_up.txt")

#######
Idents(Validation) <- "GDO_comb"
pseudo_ifnb <- AggregateExpression(Validation,assays = "RNA",return.seurat = T)

##########
pseudo_sub <- subset(pseudo_ifnb,idents = c("PSMD13","NOL9","ETV4","MIOS","THG1L",
"TFRC","ATRIP","DNAJC19","VIM","MAPRE3",
"FUNDC2","FOXN1","SEC61G","ADAM10","RPS6KA1",
"TAF1","TIMM23B","PLCG1","WDTC1","HAUS2","NR2E1","SIRT7","CD55"),invert = T)
DefaultAssay(pseudo_sub) <- "RNA"
pseudo_sub <- NormalizeData(pseudo_sub)
features2 <- rownames(
  GetAssayData(pseudo_sub, assay = "RNA", layer = "data")
)
pseudo_sub <- ScaleData(pseudo_sub,assay = "RNA",features = features2)

psuedo_mat <- GetAssayData(pseudo_sub,assay = "RNA",layer = "scale.data")
mat_subset <- psuedo_mat[rownames(psuedo_mat) %in% c(PERK_up$V1,ATF6_UPR$V1,IRE1a_UPR$V1,TSC1_DE$gene,MTOR_DE$gene,MED_DE$gene), ]

set.seed(1)
k_samples <- 7  # Adjust as needed
kmeans_samples <- kmeans(t(mat_subset), centers = k_samples, nstart = 25)
sample_clusters <- kmeans_samples$cluster

set.seed(1)
k_genes <- 7  # Adjust as needed
kmeans_genes <- kmeans(mat_subset, centers = k_genes, nstart = 25)
gene_clusters <- kmeans_genes$cluster

# Order genes and samples
gene_order <- order(gene_clusters)
sample_order <- order(sample_clusters)

expr_ordered <- mat_subset[gene_order, sample_order]
gene_clusters_ordered <- gene_clusters[gene_order]
sample_clusters_ordered <- sample_clusters[sample_order]

library(ComplexHeatmap)
library(RColorBrewer)

# Color palettes
gene_cluster_colors <- setNames(RColorBrewer::brewer.pal(k_genes, "Set3"), as.character(1:k_genes))
sample_cluster_colors <- setNames(RColorBrewer::brewer.pal(k_samples, "Set3"), as.character(1:k_samples))

# Annotations
row_anno <- rowAnnotation(Cluster = factor(gene_clusters_ordered),
                          col = list(Cluster = gene_cluster_colors))

col_anno <- HeatmapAnnotation(Cluster = factor(sample_clusters_ordered),
                              col = list(Cluster = sample_cluster_colors))


custom_sample_cluster_order <- c(5,2,4,7,3,6,1)
column_split_factor <- factor(sample_clusters_ordered,
                              levels = custom_sample_cluster_order)

# Convert sample cluster assignments to a data frame
sample_df <- data.frame(ID = names(sample_clusters_ordered),
                        Cluster = sample_clusters_ordered)

Validation_clust_sub <- subset(Validation,idents = c("PSMD13","NOL9","ETV4","MIOS","THG1L",
                                            "TFRC","ATRIP","DNAJC19","VIM","MAPRE3",
                                            "FUNDC2","FOXN1","SEC61G","ADAM10","RPS6KA1",
                                            "TAF1","TIMM23B","PLCG1","WDTC1","HAUS2","NR2E1","SIRT7","CD55"),invert = T)
FLEX_meta <- Validation_clust_sub@meta.data

median_data_FLEX <- FLEX_meta %>%
  group_by(GDO_comb) %>%
  summarise(
    median_pRPS6_ridge = median(pRPS6_ridge_GWPS),
  )

median_data_FLEX$ID <- median_data_FLEX$GDO_comb

rownames(median_data_FLEX) <- median_data_FLEX$ID
nt_med <- median_data_FLEX["NO","median_pRPS6_ridge"]
nt <- nt_med[[1]][1]
median_data_FLEX$delta_pRPS6 <- median_data_FLEX$median_pRPS6_ridge - nt

# Convert sample cluster assignments to a data frame
sample_df <- data.frame(ID = names(sample_clusters_ordered),
                        Cluster = sample_clusters_ordered)

# Join with your metadata
sample_df <- merge(sample_df, median_data_FLEX, by = "ID")  # assuming sample_metadata has "Sample" column

sample_df <- sample_df[order(sample_df$Cluster, -sample_df$delta_pRPS6), ]
sample_df$ID <- factor(sample_df$ID, levels = sample_df$ID)

# Convert Cluster to a factor with the custom order
sample_df$Cluster <- factor(sample_df$Cluster,
                            levels = custom_sample_cluster_order)

p4 <- ggplot(sample_df, aes(x = ID, y = delta_pRPS6)) +
  geom_bar(stat = "identity") +
  facet_grid(. ~ Cluster, scales = "free_x", space = "free") +  # keeps cluster separation
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid.major.x = element_blank(),  # removes vertical grid lines
        panel.grid.minor.x = element_blank(),axis.title = element_blank(),
        axis.text.x.bottom = element_blank(),panel.grid.minor.y = element_blank()) 

plot(p4)

# Ensure Sample column is character (in the right order)
ordered_sample_names <- as.character(sample_df$ID)

# Reorder expression matrix
expr_final <- expr_ordered[, ordered_sample_names]

sample_clusters_ordered <- sample_df$Cluster
names(sample_clusters_ordered) <- ordered_sample_names

custom_gene_cluster_order <- c(4,5,7,6,1,3,2)
row_split_factor <- factor(gene_clusters_ordered,
                           levels = custom_gene_cluster_order)


p5 <- Heatmap(expr_final,
              name = "Z-score",
              cluster_rows = FALSE,
              cluster_columns = FALSE,  # we already ordered columns manually
              row_split = row_split_factor,  # or however you split rows
              column_split = sample_clusters_ordered,     # uses your cluster facetting
              top_annotation = NULL,
              show_row_names = FALSE,
              show_column_names = TRUE,
              col = circlize::colorRamp2(c(-2, 0, 2), c("blue", "white", "red")))

plot(p4)
```

``` {r Fig6Apanel2, dpi=300, fig.width=10, fig.height=5, dev='png'}
plot(p5)
```

Now demonstrating the subset of genes that represent mTORC1 activaiton, repression and feedback

``` {r Fig6B, dpi=300, fig.width=10, fig.height=5, dev='png'}
Idents(Validation) <- "GDO_comb"
sub <- subset(Validation,idents = c("MTOR","RPTOR","RHEB","LAMTOR3","LAMTOR2","RRAGA","RRAGC","TSC1","DEPDC5","NPRL2","NPRL3"))

Idents(sub) <- "Complex"
sub_genes <- GetAssayData(sub,assay = "RNA")
sub_genes <- as.data.frame(sub_genes)
sub_genes <- as.data.frame(t(sub_genes))
sub_genes$ident <- sub$complex

# Reorder the factor levels for the x-axis
sub_genes$ident <- factor(sub_genes$ident, levels = c("MTORC1","RAGULATOR","TSC","GATOR1"))

p1 <- ggplot(sub_genes, aes(x = ident, y = BTG1, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("MTORC1" = "#F8766D", 
                               "RAGULATOR" = "#7CAE00", 
                               "TSC"= "#00BFC4",
                               "GATOR1" = "#C77CFF")) +
  theme_minimal() + theme(axis.title = element_blank()) + NoLegend()

p2 <- ggplot(sub_genes, aes(x = ident, y = SOCS2, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("MTORC1" = "#F8766D", 
                               "RAGULATOR" = "#7CAE00", 
                               "TSC"= "#00BFC4",
                               "GATOR1" = "#C77CFF")) +
  theme_minimal() + theme(axis.title = element_blank()) + NoLegend()

p3 <- ggplot(sub_genes, aes(x = ident, y = MAP1B, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("MTORC1" = "#F8766D", 
                               "RAGULATOR" = "#7CAE00", 
                               "TSC"= "#00BFC4",
                               "GATOR1" = "#C77CFF")) +
  theme_minimal() + theme(axis.title = element_blank()) + NoLegend()

p4 <- ggplot(sub_genes, aes(x = ident, y = IDH1, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("MTORC1" = "#F8766D", 
                               "RAGULATOR" = "#7CAE00", 
                               "TSC"= "#00BFC4",
                               "GATOR1" = "#C77CFF")) +
  theme_minimal() + theme(axis.title = element_blank()) + NoLegend()


p5 <- ggplot(sub_genes, aes(x = ident, y = ALDH1A1, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("MTORC1" = "#F8766D", 
                               "RAGULATOR" = "#7CAE00", 
                               "TSC"= "#00BFC4",
                               "GATOR1" = "#C77CFF")) +
  theme_minimal() + theme(axis.title = element_blank()) + NoLegend()

p6 <- ggplot(sub_genes, aes(x = ident, y = LCP1, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("MTORC1" = "#F8766D", 
                               "RAGULATOR" = "#7CAE00", 
                               "TSC"= "#00BFC4",
                               "GATOR1" = "#C77CFF")) +
  theme_minimal() + theme(axis.title = element_blank()) + NoLegend()


plot_grid(p1,p2,p3,p4,p5,p6)

genes_to_test <- c(
  "BTG1",
  "SOCS2",
  "MAP1B",
  "IDH1",
  "ALDH1A1",
  "LCP1"
)

pairwise_table <- do.call(rbind, lapply(genes_to_test, function(gene) {
  
  pairwise_results <- pairwise.t.test(
    x = sub_genes[[gene]],
    g = sub_genes$ident,
    p.adjust.method = "BH",
    pool.sd = FALSE
  )
  
  result_df <- as.data.frame(as.table(pairwise_results$p.value))
  
  colnames(result_df) <- c("group1", "group2", "p_adj")
  
  result_df$gene <- gene
  
  result_df <- result_df[!is.na(result_df$p_adj), ]
  
  result_df <- result_df[, c("gene", "group1", "group2", "p_adj")]
  
  return(result_df)
}))

pairwise_table


```

Exploring the dynamics of the cell cycle on the biology

``` {r Fig6D, dpi=300, fig.width=10, fig.height=5, dev='png'}
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes

Validation <- CellCycleScoring(
  object = Validation,
  s.features = s.genes,
  g2m.features = g2m.genes,
  set.ident = FALSE
)

summary <- Validation@meta.data %>%
  group_by(GDO_comb) %>%
  summarise(
    Cluster6 = median(Cluster6, na.rm = TRUE),
    G2M_proportion = mean(Phase == "G2M", na.rm = TRUE),
    n_cells = n(),
    .groups = "drop"
  )

ggplot(summary, aes(x = Cluster6, y = G2M_proportion)) +
  # Add thicker zero lines
  geom_point(color = "black", size = 3) +# Thicker vertical line# Plot median points
  geom_text_repel(aes(label = ifelse(GDO_comb %in% c("MED12","MED19","MED30","PSMD4","ASCC3","GPN3"), GDO_comb, NA)),
                  color = "red", size = 4,
                  box.padding = 0.5, max.overlaps = Inf
  ) +
  labs(
    title = "Anabolism vs G2M",
    x = "Anabolism",
    y = "% Cells G2M"
  ) +
  theme_minimal()

```


``` {r FigS6A, dpi=300, fig.width=10, fig.height=5, dev='png'}
DimPlot(Validation,group.by = "Phase")
```

``` {r FigS6B, dpi=300, fig.width=10, fig.height=5, dev='png'}
Validation_metadata <- Validation@meta.data

data_subset <- Validation_metadata %>%
  filter(GDO_comb %in% c("AARS","NARS","MED12","MED19","MED30","NO","ASCC3","GPN3","PSMD4","MED12","MED19","MED30","TSC1","DEPDC5","NPRL2","NPRL3","MTOR","RPTOR","NO","EEF1G","EIF2B3","EIF2B5","EIF3H","EIF3D","TTI1","KDM1A","GFI1B","RACK1")) %>%
  count(GDO_comb, Phase, name = "n") %>%
  group_by(GDO_comb) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()


p1 <- ggplot(data_subset, aes(x = GDO_comb, y = prop, fill = Phase)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(values = c("G1" = "#F8766D", "S" = "#619CFF", "G2M" = "#00BA38")) +
  labs(title = "Stacked Proportional Barplot",
       x = "Ident",
       y = "Proportion",
       fill = "Phase") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.title = element_blank()) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

plot(p1)
```

``` {r FigS6C, dpi=300, fig.width=10, fig.height=5, dev='png'}
cell_cycsub <- subset(Validation,idents = c("GFI1B","KDM1A","MED12","MED19","MED30","TSC1","DEPDC5","NPRL2","NPRL3","MTOR","RPTOR","NO","EEF1G","EIF2B3","EIF2B5","EIF3D","LAMTOR3","LAMTOR2","RRAGA","RRAGC","HSPE1","AARS","NARS"))

cell_cycsub_md <- cell_cycsub@meta.data     
cell_cycsub_md$ident <- cell_cycsub_md$GDO_comb

# Reorder the factor levels for the x-axis
cell_cycsub_md$ident <- factor(cell_cycsub_md$ident, levels = c("TSC1","DEPDC5","NPRL3","NPRL2","HSPE1","NO","KDM1A","EIF3D","EIF2B3","NARS","EIF2B5","EEF1G","GFI1B","MED19","RRAGA","AARS","LAMTOR2","RRAGC","MED30","MED12","LAMTOR3","MTOR","RPTOR"))

Idents(cell_cycsub) <- "Phase"
G1_ccsub <- subset(cell_cycsub, idents = "G1")
G2S_ccsub <- subset(cell_cycsub, idents = c("S","G2M"))

G1_ccsub_md <- G1_ccsub@meta.data     
G1_ccsub_md$ident <- G1_ccsub_md$GDO_comb

# Reorder the factor levels for the x-axis
G1_ccsub_md$ident <- factor(G1_ccsub_md$ident, levels = c("TSC1","DEPDC5","NPRL3","NPRL2","HSPE1","NO","KDM1A","EIF3D","EIF2B3","NARS","EIF2B5","EEF1G","GFI1B","MED19","RRAGA","AARS","LAMTOR2","RRAGC","MED30","MED12","LAMTOR3","MTOR","RPTOR"))

# Get default ggplot2 palette for all levels
default_cols <- scale_fill_discrete()$palette(length(unique(G1_ccsub_md$ident)))

# Name them after the factor levels
names(default_cols) <- levels(G1_ccsub_md$ident)

# Override only NO-TARGET color
default_cols["NO"] <- "grey"

# Apply
p1 <- ggplot(G1_ccsub_md, aes(x = ident, y = pRPS6_measured, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())

plot(p1)
```

``` {r FigS6Cb, dpi=300, fig.width=10, fig.height=5, dev='png'}
G2S_ccsub_md <- G2S_ccsub@meta.data     
G2S_ccsub_md$ident <- G2S_ccsub_md$GDO_comb

# Reorder the factor levels for the x-axis
G2S_ccsub_md$ident <- factor(G2S_ccsub_md$ident, levels = c("TSC1","DEPDC5","NPRL3","NPRL2","HSPE1","NO","KDM1A","EIF3D","EIF2B3","NARS","EIF2B5","EEF1G","GFI1B","MED19","RRAGA","AARS","LAMTOR2","RRAGC","MED30","MED12","LAMTOR3","MTOR","RPTOR"))

# Get default ggplot2 palette for all levels
default_cols <- scale_fill_discrete()$palette(length(unique(G2S_ccsub_md$ident)))

# Name them after the factor levels
names(default_cols) <- levels(G2S_ccsub_md$ident)

# Override only NO-TARGET color
default_cols["NO"] <- "grey"

# Apply
p2 <- ggplot(G2S_ccsub_md, aes(x = ident, y = pRPS6_measured, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank())

plot(p2)
```

``` {r Fig6E, dpi=300, fig.width=10, fig.height=5, dev='png'}
data_subset <- subset(df,GDO_comb %in% c("MED12","MED19","MED30","PSMD4","ASCC3","GPN3","NO"))

p1 <- ggplot(data_subset, aes(x = GDO_comb, y = prop, fill = Phase)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(values = c("G1" = "#F8766D", "S" = "#619CFF", "G2M" = "#00BA38")) +
  labs(title = "Stacked Proportional Barplot",
       x = "Ident",
       y = "Proportion",
       fill = "Phase") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.title = element_blank()) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

plot(p1)
```


``` {r Fig6F, dpi=300, fig.width=10, fig.height=5, dev='png'}
Idents(Validation) <- "Phase"
G1_sub <- subset(Validation, idents = "G1")
S_sub <- subset(Validation, idents = "S")
G2M_sub <- subset(Validation, idents = "G2M")

Idents(G1_sub) <- "GDO_comb"
Idents(G2M_sub) <- "GDO_comb"
Idents(S_sub) <- "GDO_comb"
G1_sub_F <- subset(G1_sub,idents = c("NO","TSC1","RPTOR","MED12","MED19"))
S_sub_F <- subset(S_sub,idents = c("NO","TSC1","RPTOR","MED12","MED19"))
G2M_sub_F <- subset(G2M_sub,idents = c("NO","TSC1","RPTOR","MED12","MED19"))
levels(G1_sub_F) <- c("TSC1","NO","RPTOR","MED12","MED19")
levels(S_sub_F) <- c("TSC1","NO","RPTOR","MED12","MED19")
levels(G2M_sub_F) <- c("TSC1","NO","RPTOR","MED12","MED19")

G1_sub_F_pRPS6 <- GetAssayData(G1_sub_F,assay = "ADT")
G1_sub_F_pRPS6 <- as.data.frame(G1_sub_F_pRPS6)
G1_sub_F_pRPS6 <- as.data.frame(t(G1_sub_F_pRPS6))
G1_sub_F_md <- G1_sub_F@meta.data     
G1_sub_F_pRPS6$ident <- G1_sub_F_md$GDO_comb
G1_sub_F_pRPS6$ident <- factor(G1_sub_F_pRPS6$ident, levels = c("TSC1","NO","RPTOR","MED12","MED19"))
p1 <- ggplot(G1_sub_F_pRPS6, aes(x = ident, y = pRPS6, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("TSC1" = "#F8766D", 
                               "NO" = "grey", 
                               "RPTOR"= "#F8766D",
                               "MED12" = "#F8766D", 
                               "MED19"= "#F8766D")) +
  theme_minimal() + NoLegend() + theme(axis.title = element_blank(),axis.text.x = element_text(angle = 45, hjust = 1)) + ylim(1.5,5.5)

S_sub_F_pRPS6 <- GetAssayData(S_sub_F,assay = "ADT")
S_sub_F_pRPS6 <- as.data.frame(S_sub_F_pRPS6)
S_sub_F_pRPS6 <- as.data.frame(t(S_sub_F_pRPS6))
S_sub_F_md <- S_sub_F@meta.data     
S_sub_F_pRPS6$ident <- S_sub_F_md$GDO_comb
S_sub_F_pRPS6$ident <- factor(S_sub_F_pRPS6$ident, levels = c("TSC1","NO","RPTOR","MED12","MED19"))
p2 <- ggplot(S_sub_F_pRPS6, aes(x = ident, y = pRPS6, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("TSC1" = "#619CFF", 
                               "NO" = "grey", 
                               "RPTOR"= "#619CFF",
                               "MED12" = "#619CFF", 
                               "MED19"= "#619CFF")) +
  theme_minimal() + NoLegend() + theme(axis.title = element_blank(),axis.text.x = element_text(angle = 45, hjust = 1)) + ylim(1.5,5.5)

G2M_sub_F_pRPS6 <- GetAssayData(G2M_sub_F,assay = "ADT")
G2M_sub_F_pRPS6 <- as.data.frame(G2M_sub_F_pRPS6)
G2M_sub_F_pRPS6 <- as.data.frame(t(G2M_sub_F_pRPS6))
G2M_sub_F_md <- G2M_sub_F@meta.data     
G2M_sub_F_pRPS6$ident <- G2M_sub_F_md$GDO_comb
G2M_sub_F_pRPS6$ident <- factor(G2M_sub_F_pRPS6$ident, levels = c("TSC1","NO","RPTOR","MED12","MED19"))
p3 <- ggplot(G2M_sub_F_pRPS6, aes(x = ident, y = pRPS6, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = c("TSC1" = "#00BA38", 
                               "NO" = "grey", 
                               "RPTOR"= "#00BA38",
                               "MED12" = "#00BA38", 
                               "MED19"= "#00BA38")) +
  theme_minimal() + NoLegend() + theme(axis.title = element_blank(),axis.text.x = element_text(angle = 45, hjust = 1)) + ylim(1.5,5.5)


plot_grid(p1,p2,p3,ncol =3)
```

plotting the qPCR data using the values from extended data

``` {r Fig6G, dpi=300, fig.width=10, fig.height=5, dev='png'}
qPCR_results <- read_csv("/brahms/blairj/FLEX/Validation/qPCR_results_box.csv")
p1 <- ggplot(qPCR_results, aes(x = Targ_Treat, y = sXBP1_tXBP1, fill = Target)) +
  geom_boxplot() +
  scale_fill_manual(values = c(
    "NT" = "grey70",
    "PPA1" = "#F8766D",
    "DHDDS" = "#7CAE00"
  )) + 
  theme_classic()
plot(p1)

PPA1_C_p <- t.test(
  sXBP1_tXBP1 ~ Targ_Treat,
  data = subset(qPCR_results, Targ_Treat %in% c("PPA1_C", "NT_C"))
)

NT_T_p <- t.test(
  sXBP1_tXBP1 ~ Targ_Treat,
  data = subset(qPCR_results, Targ_Treat %in% c("NT_T", "NT_C"))
)

DHDDS_C_p <- t.test(
  sXBP1_tXBP1 ~ Targ_Treat,
  data = subset(qPCR_results, Targ_Treat %in% c("DHDDS_C", "NT_C"))
)
```

Plotting the western blot data using the values from extended data

``` {r Fig6I, dpi=300, fig.width=10, fig.height=5, dev='png'}
# Create data frame
df <- data.frame(
  NT = c(1.567848014, 0.560132715, 0.917959278, 1.086919924),
  AARS = c(1.739789961, 1.220771823, 3.355173509, 2.860957359),
  RACK1 = c(2.490683122, 1.878233777, 4.252073806, 4.169747026)
)

t.test(df$NT,df$AARS)
t.test(df$NT,df$RACK1)

df_long <- df %>%
  pivot_longer(cols = everything(),
               names_to = "Condition",
               values_to = "Value")

df_long$Condition <- factor(
  df_long$Condition,
  levels = c("NT", "AARS", "RACK1")
)

p1 <- ggplot(df_long, aes(x = Condition, y = Value)) +
  geom_boxplot(width = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 2) +
  theme_classic() +
  labs(
    title = "ATF4",
    x = "",
    y = "Value"
  )

df <- data.frame(
  NT = c(1.018884253,
         0.978960573,
         0.880592662,
         1.156663332),
  AARS = c(0.042329873,
           0.035444008,
           0.454965047,
           0.467437116),
  RACK1 = c(0.040730594,
            0.040835197,
            0.943316021,
            0.906179789)
)

t.test(df$NT,df$AARS)
t.test(df$NT,df$RACK1)

df_long <- df %>%
  pivot_longer(cols = everything(),
               names_to = "Condition",
               values_to = "Value")

df_long$Condition <- factor(
  df_long$Condition,
  levels = c("NT", "AARS", "RACK1")
)

p2 <- ggplot(df_long, aes(x = Condition, y = Value)) +
  geom_boxplot(width = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 2) +
  theme_classic() +
  labs(
    title = "pRPS6",
    x = "",
    y = "Value"
  )

####
df <- data.frame(
  NT = c(1.001832864,
         0.997594519,
         0.744233601,
         1.271700492),
  AARS = c(0.398828301,
           0.503123089,
           1.557703243,
           1.175021489),
  RACK1 = c(0.268502595,
            0.311299039,
            2.524368263,
            2.430656195)
)

df_long <- df %>%
  pivot_longer(cols = everything(),
               names_to = "Condition",
               values_to = "Value")

df_long$Condition <- factor(
  df_long$Condition,
  levels = c("NT", "AARS", "RACK1")
)

p3 <- ggplot(df_long, aes(x = Condition, y = Value)) +
  geom_boxplot(width = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 2) +
  theme_classic() +
  labs(
    title = "p4EBP1",
    x = "",
    y = "Value"
  )


plot_grid(p1,p2,p3,ncol =1)
```

```{r FigS6D, dpi=300, fig.width=10, fig.height=5, dev='png'}
#GWPS data
data_subset <- GWPS_metadata %>%
  filter(gene %in% c("TSC1", "non-targeting", "RPTOR", "MED12")) %>%
  count(gene, Phase, name = "n") %>%
  group_by(gene) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()


p1 <- ggplot(data_subset, aes(x = gene, y = prop, fill = Phase)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(values = c("G1" = "#F8766D", "S" = "#619CFF", "G2M" = "#00BA38")) +
  labs(title = "Stacked Proportional Barplot",
       x = "Ident",
       y = "Proportion",
       fill = "Phase") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.title = element_blank()) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )


#uses the cell cycle percentrage data from FigS7
data_subset <- read_csv("CellCycle_dfsubset.csv")
order_vec <- c("TSC1","NO-TARGET", "RPTOR", "MED12")
data_subset$Var1 <- factor(data_subset$Var1, levels = order_vec)

p2 <- ggplot(data_subset, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(values = c("G1" = "#F8766D", "S" = "#619CFF", "G2M" = "#00BA38")) +
  labs(title = "Stacked Proportional Barplot",
       x = "Ident",
       y = "Proportion",
       fill = "Phase") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.title = element_blank()) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )


plot_grid(p1,p2)
```


```{r FigS6E, dpi=300, fig.width=10, fig.height=5, dev='png'}
pRPS6_G1 <- read_csv("pRPS6_G1.csv")

df_long <- pivot_longer(pRPS6_G1,
                        cols = everything(),
                        names_to = "Gene",
                        values_to = "Value")

df_long$log_value <- log10(df_long$Value + 1)

df_long$Gene <- factor(
  df_long$Gene,
  levels = c("TSC1","NT","RPTOR","MED12")
)

p1 <- ggplot(df_long, aes(x = Gene, y = log_value, fill = Gene)) +
  geom_boxplot() +
  scale_fill_manual(values = c(
    "NT" = "grey70",
    "MED12" = "#F8766D",
    "RPTOR" = "#F8766D",
    "TSC1" = "#F8766D"
  )) + 
  theme_classic() + 
  coord_cartesian(ylim = c(2.6,2.9)) + theme(legend.position = "none")

t.test(pRPS6_G1$NT,pRPS6_G1$MED12)
t.test(pRPS6_G1$NT,pRPS6_G1$RPTOR)
t.test(pRPS6_G1$NT,pRPS6_G1$TSC1)


#######
pRPS6_G2M <- read_csv("pRPS6_G2M.csv")

df_long <- pivot_longer(pRPS6_G2M,
                        cols = everything(),
                        names_to = "Gene",
                        values_to = "Value")

df_long$log_value <- log10(df_long$Value + 1)

df_long$Gene <- factor(
  df_long$Gene,
  levels = c("TSC1","NT","RPTOR","MED12")
)

p2 <- ggplot(df_long, aes(x = Gene, y = log_value, fill = Gene)) +
  geom_boxplot() +
  scale_fill_manual(values = c(
    "NT" = "grey70",
    "MED12" = "#00BA38",
    "RPTOR" = "#00BA38",
    "TSC1" = "#00BA38"
  )) + 
  theme_classic() + 
  coord_cartesian(ylim = c(2.6,2.9)) + theme(legend.position = "none")

t.test(pRPS6_G2M$NT,pRPS6_G2M$MED12)
t.test(pRPS6_G2M$NT,pRPS6_G2M$RPTOR)
t.test(pRPS6_G2M$NT,pRPS6_G2M$TSC1)



#######
pRPS6_S <- read_csv("pRPS6_S.csv")

df_long <- pivot_longer(pRPS6_S,
                        cols = everything(),
                        names_to = "Gene",
                        values_to = "Value")

df_long$log_value <- log10(df_long$Value + 1)

df_long$Gene <- factor(
  df_long$Gene,
  levels = c("TSC1","NT","RPTOR","MED12")
)

p3 <- ggplot(df_long, aes(x = Gene, y = log_value, fill = Gene)) +
  geom_boxplot() +
  scale_fill_manual(values = c(
    "NT" = "grey70",
    "MED12" = "#619CFF",
    "RPTOR" = "#619CFF",
    "TSC1" = "#619CFF"
  )) + 
  theme_classic() + 
  coord_cartesian(ylim = c(2.6,2.9)) + theme(legend.position = "none")

t.test(pRPS6_S$NT,pRPS6_S$MED12)
t.test(pRPS6_S$NT,pRPS6_S$RPTOR)
t.test(pRPS6_S$NT,pRPS6_S$TSC1)

plot_grid(p1,p3,p2,ncol =3)
```

Checking out RNA changes in a subset of perturbations in different clusters

``` {r FigS6F, dpi=300, fig.width=10, fig.height=5, dev='png'}
f8_PPP <- subset(Validation,idents = c("PSMD4","ASCC3","GPN3","NO"))
f8_PPP_rna <- GetAssayData(f8_PPP,assay = "RNA")
f8_PPP_rna <- as.data.frame(f8_PPP_rna)
f8_PPP_rna <- as.data.frame(t(f8_PPP_rna))
f8_PPP_rna_md <- f8_PPP@meta.data     
f8_PPP_rna$ident <- f8_PPP_rna_md$GDO_comb

# Reorder the factor levels for the x-axis
f8_PPP_rna$ident <- factor(f8_PPP_rna$ident, levels = c("ASCC3","GPN3","PSMD4","NO"))

# Get default ggplot2 palette for all levels
default_cols <- scale_fill_discrete()$palette(length(unique(f8_PPP_rna$ident)))

# Name them after the factor levels
names(default_cols) <- levels(f8_PPP_rna$ident)

# Override only NO-TARGET color
default_cols["NO"] <- "grey"

p1 <- ggplot(f8_PPP_rna, aes(x = ident, y = DDIT3, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

p2 <- ggplot(f8_PPP_rna, aes(x = ident, y = PPP1R15A, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

png("plots/FigS6G_PPP_rna_BoxPlot_.png",width = 2400, height = 4000,res=600)
plot_grid(p1,p2,ncol =1)
dev.off()


control_vals <- f8_PPP_rna$DDIT3[f8_UPR_rna$ident == "NO"]
mean_control <- mean(control_vals)
results <- f8_PPP_rna %>%
  filter(ident != "NO") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(DDIT3),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(DDIT3, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results

control_vals <- f8_PPP_rna$PPP1R15A[f8_PPP_rna$ident == "NO"]
mean_control <- mean(control_vals)
results <- f8_PPP_rna %>%
  filter(ident != "NO") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(PPP1R15A),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(PPP1R15A, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results
```

``` {r FigS6G, dpi=300, fig.width=10, fig.height=5, dev='png'}
f8_UPR <- subset(Validation,idents = c("PPA1","SRP72","XRCC6","NO"))
f8_UPR_rna <- GetAssayData(f8_UPR,assay = "RNA",slot = "data")
f8_UPR_rna <- as.data.frame(f8_UPR_rna)
f8_UPR_rna <- as.data.frame(t(f8_UPR_rna))
f8_UPR_rna_md <- f8_UPR@meta.data     
f8_UPR_rna$ident <- f8_UPR_rna_md$GDO_comb

# Reorder the factor levels for the x-axis
f8_UPR_rna$ident <- factor(f8_UPR_rna$ident, levels = c("PPA1","SRP72","XRCC6","NO"))

# Get default ggplot2 palette for all levels
default_cols <- scale_fill_discrete()$palette(length(unique(f8_UPR_rna$ident)))

# Name them after the factor levels
names(default_cols) <- levels(f8_UPR_rna$ident)

# Override only NO-TARGET color
default_cols["NO"] <- "grey"

p1 <- ggplot(f8_UPR_rna, aes(x = ident, y = SSR2, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

p2 <- ggplot(f8_UPR_rna, aes(x = ident, y = SEC61A1, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

plot_grid(p1,p2,ncol =1)

control_vals <- f8_UPR_rna$SSR2[f8_UPR_rna$ident == "NO"]
mean_control <- mean(control_vals)
results <- f8_UPR_rna %>%
  filter(ident != "NO") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(SSR2),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(SSR2, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results

control_vals <- f8_UPR_rna$SEC61A1[f8_UPR_rna$ident == "NO"]
mean_control <- mean(control_vals)
results <- f8_UPR_rna %>%
  filter(ident != "NO") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(SEC61A1),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(SEC61A1, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results
```

``` {r FigS6H, dpi=300, fig.width=10, fig.height=5, dev='png'}
Idents(Validation) <- "GDO_comb"
sub2 <- subset(Validation, idents = c("EIF2B3","EIF2B5","AARS","EEF1G","NARS","NARS","NO"))

sub2_rna <- GetAssayData(sub2,assay = "RNA")
sub2_rna <- as.data.frame(sub2_rna)
sub2_rna <- as.data.frame(t(sub2_rna))
sub2_md <- sub2@meta.data     
sub2_rna$ident <- sub2_md$GDO_comb

# Reorder the factor levels for the x-axis
sub2_rna$ident <- factor(sub2_rna$ident, levels = c("NO","NARS","AARS","EEF1G","EIF2B3","EIF2B5"))

# Get default ggplot2 palette for all levels
default_cols <- scale_fill_discrete()$palette(length(unique(sub2_rna$ident)))

# Name them after the factor levels
names(default_cols) <- levels(sub2_rna$ident)

# Override only NO-TARGET color
default_cols["NO"] <- "grey"

p1 <- ggplot(sub2_rna, aes(x = ident, y = PHGDH, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

p2 <- ggplot(sub2_rna, aes(x = ident, y = EIF1, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())


control_vals <- sub2_rna$SESN2[sub2_rna$ident == "NO"]
mean_control <- mean(control_vals)
results <- sub2_rna %>%
  filter(ident != "NO") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(SESN2),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(SESN2, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results

control_vals <- sub2_rna$DDIT4[sub2_rna$ident == "NO"]
mean_control <- mean(control_vals)
results <- sub2_rna %>%
  filter(ident != "NO") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(DDIT4),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(DDIT4, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results

plot_grid(p1,p2)
```

``` {r FigS6I, dpi=300, fig.width=10, fig.height=5, dev='png'}
sub2_pRPS6 <- GetAssayData(sub2,assay = "ADT")
sub2_pRPS6 <- as.data.frame(sub2_pRPS6)
sub2_pRPS6 <- as.data.frame(t(sub2_pRPS6))
sub2_md <- sub2@meta.data     
sub2_pRPS6$ident <- sub2_md$GDO_comb

# Reorder the factor levels for the x-axis
sub2_pRPS6$ident <- factor(sub2_pRPS6$ident, levels = c("NO","NARS","AARS","EEF1G","EIF2B3","EIF2B5"))

# Get default ggplot2 palette for all levels
default_cols <- scale_fill_discrete()$palette(length(unique(sub2_pRPS6$ident)))

# Name them after the factor levels
names(default_cols) <- levels(sub2_pRPS6$ident)

# Override only NO-TARGET color
default_cols["NO"] <- "grey"

p1 <- ggplot(sub2_pRPS6, aes(x = ident, y = pRPS6, fill = ident)) +
  geom_boxplot() +
  scale_fill_manual(values = default_cols) +
  theme_minimal() +
  ylim(2, 5.5) +
  NoLegend() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

control_vals <- sub2_pRPS6$pRPS6[sub2_pRPS6$ident == "NO"]
mean_control <- mean(control_vals)
results <- sub2_pRPS6 %>%
  filter(ident != "NO") %>%
  group_by(ident) %>%
  summarise(
    mean_group = mean(pRPS6),
    log2FC = log2(mean_group / mean_control),
    percent_change = (2^log2FC - 1) * 100,
    p_value = t.test(pRPS6, control_vals)$p.value,
    .groups = "drop"
  ) %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))

results

plot(p1)
```